### 2016-02-17 main.tex

% \documentclass[conference]{IEEEtran}

% \begin{document}
% \title{An Important Conference Contribution}

% \author{\IEEEauthorblockN{Author One\IEEEauthorrefmark{1},
% Author Two\IEEEauthorrefmark{2}, Author Three\IEEEauthorrefmark{3} and
% Author Four\IEEEauthorrefmark{4}}
% \IEEEauthorblockA{Department of Whatever,
% Whichever University\\
% Wherever\\
% Email: \IEEEauthorrefmark{1}author.one@add.on.net,
% \IEEEauthorrefmark{2}author.two@add.on.net,
% \IEEEauthorrefmark{3}author.three@add.on.net,
% \IEEEauthorrefmark{4}author.four@add.on.net}}
% \maketitle

% \end{document}

\documentclass[conference]{IEEEtran}
\usepackage{blindtext, graphicx}
\usepackage{fancyhdr} 
\graphicspath{ {images/} }
\fancyhf{}
\cfoot{\thepage}
\pagestyle{fancy}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\pagenumbering{arabic}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Investigating Fault-Proneness Through Hotfix Commits: A Research in the Light of Different Programming Languages}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Martin Lazarov\IEEEauthorrefmark{1},
Ismaa'eel Ataullah\IEEEauthorrefmark{2}, Vladyslav Kolesnyk\IEEEauthorrefmark{3},
Veselin Pavlov\IEEEauthorrefmark{4}}
\IEEEauthorblockA{Department of Computer Science,
UCL\\
Email: \IEEEauthorrefmark{1}martin.lazarov.12@ucl.ac.uk,
\IEEEauthorrefmark{2}ismaaeel.ataullah.12@ucl.ac.uk,
\IEEEauthorrefmark{3}vladyslav.kolesnyk.12@ucl.ac.uk,
\IEEEauthorrefmark{4}veselin.pavlov.15@ucl.ac.uk}}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Choosing the most suitable tools and programming languages is a very challenging task for every project manager and developer. Often the success of a software project is determined by the stability of the system after a release. In this paper, we propose analysing hotfix commits in various development areas as a means of indicating what languages tend to produce the least amount of problems after software has been published. We have focused on parsing metadata from projects hosted on SourceForge related to five specific topics (“software development”, “databases”, “security”, “games”, “front-ends”) and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages. 
\end{abstract}

\begin{IEEEkeywords}
Version Control, hotfixes, SourceForge, commits metadata.
\end{IEEEkeywords}

\section{Introduction}
Planning and resource allocation are arguably the most crucial components of any software project. Our research was motivated by the fact that very often, in the planning phase project managers neglect the importance of hotfixes – which becomes essential once a project is deployed. The most common reason for that practice is the fact that hotfixes are not constant and their presence is difficult for prediction. As a result, it is quite challenging to estimate their quantity and incorporate them into the cost of the project.  Our research attempts to provide this crucial information by correlating hotfixes with the choice of language to develop software in a specific area. As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: “a corrective commit deployed within 1 day, following a project’s version update to fix critical issues in the release”.\par
We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used. Consequently, project managers will be able to make a more informed decision with regard to the most appropriate programming languages they could use to try to achieve a minimum number of problems in production. This will improve software engineering practices and the overall project planning activity. Furthermore, it will help optimise the allocation of resources.
The contributions of this paper are:

\begin{itemize}
  \item parameters to determine most probable hotfix commits in the repository.
  \item a quantitative analysis identifying fault-proneness of specific programming topics and the relevant distribution of faults across programming languages.
  \item an investigation into the following relationships between languages: strongly-typed vs weakly-typed, static vs dynamic and functional vs procedural to try to determine what group of languages is the most reliable.
  \item a foundation for future research in the area of identifying language-specific hotfixes.
\end{itemize}
\par
\textbf{Structure of the paper.}\\
We discuss Related Work in section 2. The next segment discusses our Solution. Following from that, an analysis of the gathered results is presented. Finally, we review our approach and findings, and examine Future Work.

\section{Related Work}
Many researchers have explored various aspects of defects in software through the use of versioning system history logs. We have also decided to follow this approach due to its openness and accessibility. This method is also very widely adopted amongst the MSR community [1]. Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages’ stability.

\subsection{Identifying hotfix commits}
The first major problem we had to solve was identifying actual hotfix commits. Due to Boa’s limitations, we were not able to find commit history from hotfix branch. We were only capable of collecting information with regard to commits performed on the main branch. Therefore, we reviewed various papers to identify the most appropriate method of determining the type of the commit to extract only the ones related to a bug correction.\par

A large number of the reports that we studied implemented a very simple approach. They focused on the commit message as the main mechanism to try to identify whether a commit was of a corrective type [1, 4, 5]. They simply investigated if the string of text contained particular key words, such as: “bug”, “fix”, “issue”, “error”, “broke”, etc.\par

Some authors proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages. Their algorithm consisted of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g. “for”, “code”, “to”, “the”, etc. ) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits [3].\par

Another approach to identify bug fix commits involved performing a classification based on the size of the commits [1, 3, 5]. The size of a commit can be measured based on the number of lines or the amount of files that were changed. According to these authors, large commits tend to improve code quality or add new features, while small ones are considered to be corrective - targeted at fixing bugs in the project. Some researchers proposed a further fine-grained size classification depending on the number of files changed per commit [1]. They defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large (126 up ). The scientists further argued that most of the commits tend to be tiny (in 80\% of the cases). This implies that bug fix commits occur very often as corrective commits tend to be small ones.\par

We also looked at the time factor with regard to commits. According to some authors, issues in critical components (e.g. errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered [2]. In [3], the researchers argue that the time between changes will depend on the type of alteration. Furthermore, they hold the view that corrective changes require the least amount of time to be completed as often they have to be performed promptly regardless of the complexity of the task.\par

In terms of the time aspect, there was another paper which investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk [10]. However, they explored branches in general without focusing specifically on bugfix branches. Some other scientists were trying to explain people’s behaviour during commits by investigating their commit frequency distribution. As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits [11].\par

Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed [6]. However, they did not make their investigation language-specific as opposed to what we propose in our paper.\par

Other scientists tried to explore the logical coupling of commits as a way of identifying potential software defects. They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future [9].\par

A further interesting research that we found was related to the use of separate bug fix branches. According to the study performed by Phillips et al. [7], 41\% of the tested projects had incorporated an isolated bug fix branch.

\subsection{Comparison of programming languages' stability}
Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality [8]. Furthermore, there were several papers that compared and contrasted the performance of different languages, which led to our second research topic. We explored these existing scientific studies to see what techniques authors had implemented. We also analysed the hypotheses these papers proposed to increase the scope of our project.\par

Our specific point of interest was related to different software topics and programming languages to discover how number of hotfixes per topic depend on the language developers use.\par

There were researchers that focused on the effect of programming languages on software quality [12]. The authors based performance on the rate of defect occurrences in various languages. The study was carried out on a huge dataset from GitHub (1.5 million commits). To evaluate the effect that programming languages have on software quality, the authors implemented the following strategy. First, they grouped languages according to their type, i.e. strongly vs weakly typed. They also used mathematical methods to triangulate and produce statistics for how group size, project history and number of contributors affect the number of defects. In order to achieve more representative results, they took a sample of 17 languages and extracted the top 50 projects per language. The results showed how defect-prone languages are compared to others. This was done by taking a grand mean and then comparing each language to the grand mean. Languages were then grouped, a mean was taken and then types of languages were once again compared to a grand mean - e.g. functional-strong-static languages compared to functional-dynamic-strong languages. This research focused on classifying bugs based on their type as opposed to our investigation, where we focused specifically on the distribution of hotfixes across programming languages.\par

In [13] Java and C++ were evaluated by comparing bug rates. In this case, there were two definitions for bugs - a defect and a bug itself. A defect included issues such as syntax errors, logical errors and changes in variable name. A bug itself was only defined to be a problem detected during testing or deployment. The mean number of defects and bugs, averaged over a number of projects, was compared between both languages per 1000 lines of code. The scientists also compared observed defects and bugs per hour as well as the time taken to fix defects in both languages. Productivity was also measured by analysing number of lines written per second.\par

There were researchers that used slightly different parameters to compare programming languages, these included program length, programming effort, run-time efficiency, memory consumption, and reliability [14].\par

An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages [15]. This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.\par

Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness [16]. They argued that conciseness is an important factor as a more compact language can help write a program with fewer bugs.

\section{Solution}
Our hypothesis was motivated by research paper [12]. In it the authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.\par

We were interested in investigating whether their findings applied when comparing number of hotfixes to total number of commits and if those conclusions were still relevant when we divided our project sources into various topics.\par

More specifically we wanted to see whether:
\begin{enumerate}
\item \textbf{Strongly-typed languages required fewer hotfixes than weakly-typed languages.}
\item \textbf{Statically-typed languages required fewer hotfixes than dynamically-typed languages.}
\item \textbf{Functional languages required fewer hotfixes than procedural ones.}
\end{enumerate}
\par
In order to evaluate our hypothesis, we developed a tool in Boa [17] that would allow us to incorporate our ideas and perform an analysis to verify them. Next, we present our reasoning and what factors we considered to accomplish our research.\par

To calculate the frequency of hotfixes for various projects and topics, we had to design a way of distinguishing a hotfix based on the commit messages. Based on the definition of a hotfix, we focused on discovering corrective commits that were submitted quickly after a new version release. In this way, we were hoping to identify problems specifically related to the version update. Therefore, we decided to divide our implementation into: detecting a \textbf{version bump commit} (i.e. a commit indicating a version increase) and identifying subsequent \textbf{bug fix commits}.\par

Unfortunately, no prior projects implemented a technique to detect version bump commits. However, from the observed data it turned out to be quite straightforward to extract this information by generating a regular expression to match strings like: “bump”, “version”, “update” in an appropriate order.\par

From our Related Work, we found projects [1, 4, 5], which distinguished bug fix commits only based on the commit messages. However, due to the similarity between commit messages for normal bug fixes and hotfix commits, and occasionally the large sequences of text, it was challenging to use the same approach. Therefore, in addition to bug fix commit detection, we applied some other parameters to ensure that it is not just a regular fix, but a hotfix. As previously discussed, we used a regular expression which examined commit messages for words associated with a corrective commit (e.g. “bug”, “fix”, “issue”, “error”, “broke”, etc.) to detect all bug fix commits. Next, we considered other possibilities that would let us detect hotfixes in a more reliable way. Some of the papers we discussed in Related Work [1, 3, 5, 11] incorporated metadata, such as: the size of changes (the number of files that were altered) and the time of the commit (how long it took for a patch to be submitted) to identify bug fixes. We considered these options for our purposes as well. Consequently, in order to increase our confidence level that the current fix commit is an actual hotfix, we introduced the following parameters to verify the emergency of the fix:
\begin{itemize}
  \item \textbf{Time} - commits have to be submitted within 1 day from the version bump commit
  \item \textbf{Look-ahead} range for commits - only n numbers of commits should be considered after version bump
\end{itemize}

\par
The size of the commit (i.e. Number of files changed ) could have been used as another parameter to determine hotfixes as our investigation discovered that corrective type of changes generally affect few files (e.g. revert new alterations or switch off a problematic feature). However, based on our results, we found examples where lots of changes were merged into one commit. Consequently, several hotfixes may be submitted as part of a single merge commit, hence affecting many files. In such case, we would fail to account for these hotfix commits because of the large number of files involved in the fix.\par

The reason why we decided to put a limit and investigate commits only within 1 day after a release was because of the nature of hotfixes. Whenever a live product contains a critical fault, it tends to be identified quickly as key functionalities stop working. Therefore, leaving bugs in production amplify the risk of the system being unusable. As a result, such hotfixes are assigned the highest priority and have to be performed rapidly even outside of the normal development process.\par

Furthermore, it is very likely that fixes performed on the first day after a new release will refer to problems occurring before the version increase (or at latest to commits submitted on that day). Therefore, we decided to focus on the Look-ahead to investigate the first n commits after a version update. We focused on the first 15 commits as anything after this point would increase the probability that a fix is related to code added after the release rather than before that. To provide more fine-grained results, we looked at the initial 5, 10 and 15 commits separately.\par

With regard to evaluating the most optimal Look-ahead value, we decided to count the number of Forward Engineering commits as well (i.e. commits that introduce new features). We did this because we were interested in identifying the largest fix commits/forward commits ratio. This would allow us to guarantee that lower number of new features were introduced into the code and consequently that some of the higher number of fixes refer to the previous version.\par

Based on our research and the experiments we ran, we managed to identify the most optimal parameters (i.e. \textbf{time}: within 1 day after a release and \textbf{look-ahead}: the first 10 commits - \textit{see Table 1 in Appendix}). Having accomplished that, we focused on selecting a dataset. We decided to analyse SourceForge 2013 because it provided categorisation of projects into programming topics (e.g. code related to: “databases”, “software development”, “security”, etc.). Unfortunately, Github did not possess such classification. We believed that being able to sort projects based on topics would produce greater insights and more useful information for developers, as they would be more interested in language efficiency for particular software development area.\par

From all the available projects, we only concentrated on the ones that had incorporated versioning (i.e. containing bump commits). Our argument was that only if there were version updates which we could detect, we would be able to perform a hotfix analysis. Therefore, we extracted all the projects that had incorporated bump commits and applied our parameters to try to recognise the hotfixes. Consequently, for each programming language we analysed the number of hotfix and bump commits so that we would be able to 
\begin{enumerate}
\item \textbf{Strongly-typed languages required fewer hotfixes than weakly-typed languages.}
\item \textbf{Statically-typed languages required fewer hotfixes than dynamically-typed languages.}
\item \textbf{Functional languages required fewer hotfixes than procedural ones.}
\end{enumerate}

\subsection{Overall performance of programming languages}
We started with investigating the overall reliability of different languages regardless of the topic they belonged to.\\
\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{Bar_chart}
\caption{Overall hotfixes per programming languages for 100 releases.}
\label{fig:my_label}
\end{figure}

\par
By performing a general analysis on all languages regardless of their topic, we compared the ratios of hotfix commit/release based on the available data. According to our results, Scala is the most reliable programming language with only 6.8 hotfixes per 100 releases ( \textit{see Table 1 in Appendix for more details}). On the other hand, Matlab is the worst performing language with 37.93 hot corrections per 100 version updates.\\

\textbf{Hypothesis 1}:\par
The strongly-typed languages we considered were: C\#, Java, Python, Scala.\par
The weakly-typed languages we identified were: C, C++, Objective C, Perl, Javascript, PHP.\par

First, Scala performs better than all weakly-typed languages. However, C\#, Java and Python do not support the hypothesis and do not outperform all weakly-typed languages. Moreover, all of them have a higher hotfix per version bump ratio than Objective C. Therefore, based on our results, we can state that it is often the case that a strongly-typed language \textbf{does not} have fewer hotfixes than a weakly-typed language and our analysis is inconclusive on this matter.\\

\textbf{Hypothesis 2}:\par
Static languages considered: C, C++, Objective C, Java, Scala.\par
Dynamic languages identified: Python, Perl, PHP, Javascript.\par

C performs better than all dynamic languages apart from Perl. 
C++ is only more reliable than PHP. In all other cases, C++ contains more hotfixes than the other dynamically-typed languages. 
Objective C is a very strong static candidate and is more reliable than all dynamic languages. Java performs slightly worse, but it still has overall dominance over the dynamically-typed languages. It has more amount of hotfixes per release only when compared with Perl. However, in the other cases, it is more efficient than all languages from the second group. Scala, as shown already is the best performer throughout all languages. Therefore, it is more reliable than all dynamic ones. From this analysis, we could say that in most of the cases, the second part of our hypothesis \textbf{is supported} by the results we obtained.\\

\textbf{Hypothesis 3}:\par
Unfortunately, from the languages that we were able to collect information about, only Scala is a member of the functional programming languages. As mentioned already Scala is the best performer amongst all evaluated languages, and as such will undoubtedly have less hotfixes than all the procedural languages (e.g. C, C++,C\#, Objective C, Java ). Thus, based purely on our results, the third part of our hypothesis agrees with the data that we collected.\par

Overall, our hypothesis agreed with the results we obtained apart from a small number of cases. However, we expect that this may be due to the different amount of releases and hotfixes identified per language. We could also conclude that Functional-Static-Strong languages are better than Procedural-Dynamic-Weak ones. Therefore, we could also argue that by using combinations of the former type of languages, developers will be able to guarantee the lowest amount of hotfixes per release. As a result, the effort needed for maintenance will be reduced and software quality will be improved.

\subsection{Topic: Databases}
We take into account reliability as we want to reach reasonable scientific conclusions based on our findings. Unfortunately, we experience limitations within the data. For example, there are languages where we have identified only 2 or 3 bump commits. Deriving results from such a small dataset is extremely unreliable and it is not likely we can reach accurate solutions with them, hence we exclude them as outliers. These languages are Zope and C\# 1.0.\par
We can see from our results that C has the lowest number of hotfix to bump commit ratio, suggesting that C is the most reliable language to program in, closely followed by Python. C++, PHP and Javascript have similar reliability according to our results but are less stable when compared to Perl or Python. Given the large number of bump commits for C++, PHP and Javascript, we can assume these results are reliable.\par
Our hypothesis had three aspects to it and we will analyse whether our results support all of them.\\

\textbf{Hypothesis 1}:\par
The strongly typed languages we consider are: Java, Python.\par
The weakly typed languages we consider are: C, C++, Perl, PHP, Javascript.\par
In this case we have a mixed series of results which suggest that this part of our hypothesis may not be entirely correct. Although Python is ranked as the second most reliable language, it is topped by C, which is a weakly-typed language. If Java had been placed third,  the hypothesis would have been better supported as we could have then treated C as an outlier. However, this is not the case. Above Java is another weakly-typed language, namely Perl. In conclusion, when considering database-specific projects, we cannot say that Strongly-typed languages required fewer hotfixes when compared to weakly-typed ones.\\

\textbf{Hypothesis 2}:\par
The statically-typed languages we consider are: C, Java, C++.\par
The dynamically-typed languages we consider are: Javascript, Python, Perl, PHP.\par

Again we have a mixed list of results. C is top and this would suggest that statically-typed languages indeed do require fewer hotfixes when compared to dynamically-typed languages. However, the other static languages, Java and C++, are in the bottom half of the table in terms of reliability. Once again, we cannot say that this part of our hypothesis is supported when comparing languages used for developing database-specific programmes.\\

\textbf{Hypothesis 3}:\par
We cannot compare this aspect of the hypothesis as no languages in this set of results are classified as Functional languages.\par
\textbf{In conclusion},no elements of our hypotheses were supported by our results. In two cases, our observations were not definitive and in the other we did not have the required results to make a comparison. We can conclude that our hypothesis is not supported when considering languages used for developing  database-related projects.

\subsection{Topic: Games}
In order to only analyse those results we deem reliable, we will exclude C\#, Javascript and Objective C on the basis that the number of bump commits analysed for them were too low.\par

We can see from our results that for programming games, Java seems to be the most reliable language to use. C is second; however, it is somewhat behind, with 24.6\% of all commits being hotfix commits. C and C++ are very similar and are only separated by 0.1\%. In last place is Python which has a hotfix rate of almost 30\% and according to our results is the least reliable language to use.\\

\textbf{Hypothesis 1}:\par
The strongly-typed languages we consider are: Java, Python.\par
The weakly-typed languages we consider are: C, C++.\par

Java is the most reliable language by far according to our results. C, C++ and Python have quite similar hotfix rates. Java is a strongly-typed language which has a definitive advantage over the others. Thus, we can say this aspect of our hypothesis is somewhat supported. It’s only partially justified because while Python, C and C++ have similar hotfix rates, Python still ranks at the bottom of our results table.\\

\textbf{Hypothesis 2}:\par

The statically-typed languages we consider are: C, Java, C++.\par
The dynamically-typed languages we consider are: Python.\par

This aspect of the hypothesis is categorically supported by our results, all statically-typed languages are ranked higher than the dynamically-typed Python in our results. We can therefore conclude that static languages required fewer hotfixes when compared to dynamic languages and hence the latter are less reliable when programming games.\\

\textbf{Hypothesis 3}:\par
We cannot compare this aspect of the hypothesis as no languages in this set of results are classed as Functional languages.\par

\textbf{In conclusion}, we can state that our overall hypothesis is somewhat supported with regard to projects labeled as Games.

\subsection{Topic: Front-ends}
Here we can see a huge difference in numbers of bump commits detected in various projects. Thus, we can argue that not many projects were observed to have a Front-end label for sufficient analysis. And many other popular Front-end languages (e.g. Javascript, C\#) were not identified.\par
Hence, after removing languages with a small number of detected projects, we ended up with \textbf{Scala} and \textbf{Java}, where the former has a slightly smaller hotfixes per bump ratio.\\

\textbf{Hypothesis 1}:\par
Both Scala and Java are considered to be strongly-typed; thus, no comparison can be made within the filtered languages. 
Regarding all the languages, we can observe that Java and Scala, overall, have more reliable statistics than weakly-typed (PHP, C, Perl), hence the hypothesis is \textbf{satisfied}.\\

\textbf{Hypothesis 2}:\par
Both Scala and Java are classified as statically-typed; therefore, no comparison can be made within the filtered languages. In terms of all the languages in the list, we can observe that most of the static ones (C++, Scala and Java) are more reliable than dynamic languages (PHP, Perl), hence the hypothesis is \textbf{satisfied}.\\

\textbf{Hypothesis 3}:\par
From the set of languages, only Scala is considered to be functional and it indeed performed better than all the others, therefore the hypothesis is \textbf{satisfied}.

\subsection{Topic: Software Development}
We filtered languages that had insignificant number of projects under the Software Development label and ended up analysing: Java, C++, Python and C languages. It turns out Java appeared to be the most stable language, where on average for 10 version updates, only 1 hotfix would be detected.\\

\textbf{Hypothesis 1}:\par
There was no enough evidence to suggest that \textbf{all} strongly-typed languages (Java, Python) performed better than the weakly-typed ones (C++, C) within the Software Development topic. For example, Python(22.22\%) has more hotfixes than C++ (19.05\%). However, the average ratio for strongly-typed languages (16.5\%) is lower than the average ratio for weakly-typed (21.63\%). Thus, we can conclude that the hypothesis is \textbf{satisfied} in general.\\

\textbf{Hypothesis 2}:\par
Not all static languages would satisfy the hypothesis, as for example, Python as a dynamic language outperforms C (defined as a statically-typed language). However, by considering the average ratio of filtered languages, we can see that static languages (Java, C++, C - 19.1\%) by far outweigh the dynamic language Python - 22.2\%.Hence, the hypothesis is \textbf{satisfied}, however \textbf{dynamic languages are not represented} well in the dataset.\\

\textbf{Hypothesis 3}:\par
No functional languages were presented in dataset. Therefore, we cannot provide any conclusive argument with regard to the third component of our hypothesis.

\subsection{Topic: Security}
Due to the low number of releases per language that we were able to identify, we cannot guarantee that the results we collected would hold in general. However, in order to be able to perform some analysis on the data and reach somewhat valid results, we decided to exclude only the outlier languages  (i.e. the ones containing the lowest number of version updates). Thus, we decided to eliminate C\# and Python from the list of evaluated languages.\par

According to our filtered data, the best performing language is C with just 1 hotfix per 20 releases on average. On the other hand, the least reliable one is C++ as it requires a hotfix for approximately every 6.5 releases.\\

\textbf{Hypothesis 1}:\par
The analysis of our results reveals that strongly-typed languages (Java) do not perform better in comparison with weakly-typed languages (C, C++, Javascript, Perl, PHP). In fact, Java only outperforms C++, but in all other cases, it requires more hotfixes than the remaining weakly-typed languages (C, Javascript, Perl, PHP). Therefore, the first part of our hypothesis does not hold for projects related to Security.\\

\textbf{Hypothesis 2}:\par
Our observations were inconclusive on whether static languages (C, C++, Java) are more reliable than dynamic ones (Javascript, Perl, PHP). Based on our results, C performed better than all dynamically-typed languages. However, Java and C++ both required more hotfixes per version update than all dynamic languages. Therefore, our analysis cannot provide a convincing argument whether static languages are more reliable than dynamic ones. As a result, the second part of our hypothesis is inconclusive on this matter.\\

\textbf{Hypothesis 3}:\par
Unfortunately, from the results that we obtained, we could not find any functional languages to compare them to procedural ones (C, C++, Java).
Consequently, regarding the languages used in Security related projects, we are not capable of deriving any conclusions about the required number of hotfixes whatsoever. Thus, there is no argument to support or disprove the third element of our hypothesis.\par
Overall, due to the insufficient number of releases in our data, we were not able to state conclusively whether our hypothesis is supported or not. Although there were examples of disproving it, we believe that having information about a larger amount of releases may actually reveal results closer to the overall observations.

\subsection{Limitations.}
There were several limitations that we encountered throughout our research:
\begin{itemize}
  \item we were able to analyse only a small part of all available projects because most of the developers did not support a versioning system, i.e. no bump commits could be identified.
  \item we could not extract hotfix branches, only the overall commits on the trunk were available and thus we used our own parameters to try to distinguish hotfix commits. Therefore, we may have introduced some false positives or failed to account for actual hotfixes.
  \item the way we detected bump commits was by writing a regular expression to match words like:  “bump”, “version”, “update” in an appropriate order. Thus, we may have excluded version updates had they been indicated differently. 
  \item most open source projects are not under the stress of releasing as opposed to commercial projects. Consequently, they do not release very often and/or do not support versioning.
  \item only one dataset was used to evaluate our hypothesis - SourceForge 2013.
\end{itemize}

\section{Conclusions and Future Work}
We have presented a research in detecting the fault-proneness of programming languages through hotfix commits - a topic that has not been investigated so far. To increase the significance of our study we performed both an overall estimation of hotfixes and a topic-specific analysis. The tool that we created was written in Boa and mined repositories from SourceForge 2013’s dataset. Our goal was to verify whether the general relationships between languages’ reliability holds with regard to hotfixes. Specifically, we investigated the following relations: strongly-typed vs weakly-typed languages, static vs dynamic and functional vs procedural. We expected that in each pair the first group of languages will be dominant. Although our overall results supported our hypothesis, on a topic-by-topic basis we were not able to verify it.\par
Therefore, we believe that some future work has to be accomplished to increase the validity of our observations. In particular, more datasets could be used, including Github (provided that projects are categorised into topics). Further improvements may be done to identify bump commits more accurately. Recently, a few commercial projects have been open-sourced and as a result, we could use them to enhance the quality of our data. In such projects, releases will have occurred more frequently and thus, there is a higher probability of identifying more hotfixes.\par
Finally, we believe that our paper provides a good foundation for future research to be done in this topic. We expect that the results we obtained will help developers and project managers improve their planning and understand what the most appropriate languages are for their specific topic.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}
\bibitem{}
Hattori, L.P.; Lanza, M., \emph{"On the nature of commits"} in Automated Software Engineering - Workshops, 2008. ASE Workshops 2008. 23rd IEEE/ACM International Conference on , vol., no., pp.63-71, 15-16 Sept. 2008
\bibitem{}
Livshits,B.; Zimmerman,T.; \emph{"DynaMine: Finding Common Error Patterns by Mining Software Revision Histories"} in Proceedings of the 10th European software engineering conference held jointly with 13th ACM SIGSOFT international symposium on Foundations of software engineering, pp.296-305
\bibitem{}
Mockus, A.; Votta, L.G., \emph{"Identifying reasons for software changes using historic databases"} in Software Maintenance, 2000. Proceedings. International Conference on , vol., no., pp.120-130, 2000
\bibitem{}
Ratzinger,J.; Sigmund,T.; Gall,H.; \emph{"On the Relation of Refactoring and Software Defects"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp. 35-38
\bibitem{}
Hindle,A.; German,D.; Holt,R.; \emph{"What do large commits tell us?: a taxonomical study of large commits"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.99-108
\bibitem{}
Oliva,G.; Steinmacher,I.; Wiese,I.; Aurelio,M.; \emph{"What Can Commit Metadata Tell Us About Design Degradation?"} in IWPSE 2013, Proceedings of the 2013 International Workshop on Principles of Software Evolution, pp.18-27
\bibitem{}
Phillips,S.; Sillito,J.; Walker,R.; \emph{"Branching and merging: an investigation into current version control practices"} in CHASE '11 Proceedings of the 4th International Workshop on Cooperative and Human Aspects of Software Engineering, pp.9-15
\bibitem{}
Agrawal, K.; Amreen, S.; Mockus, A., \emph{"Commit Quality in Five High Performance Computing Projects"} in Software Engineering for High Performance Computing in Science (SE4HPCS), 2015 IEEE/ACM 1st International Workshop on , vol., no., pp.24-29, 18-18 May 2015
\bibitem{}
Steff, M.; Russo, B., \emph{"Co-evolution of logical couplings and commits for defect estimation"} in Mining Software Repositories (MSR), 2012 9th IEEE Working Conference on , vol., no., pp.213-216, 2-3 June 2012
\bibitem{}
Williams,C.; Spacco,J.; \emph{"Branching and Merging in the Repository"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.19-22
\bibitem{}
Kolassa,C.; Riehle,D.; Salim,M.; \emph{"The Empirical Commit Frequency Distribution of Open Source Projects"} in WikiSym '13, Proceedings of the 9th International Symposium on Open Collaboration, Article No.18
\bibitem{}
Ray,B.; Posnett,D.; Filkov,V.; Devanbu,P;  \emph{A Large Scale Study of Programming Languages and Code Quality"} in Github;  FSE 2014, Proceedings of the 22nd ACM SIGSOFT International Symposium on Foundations of Software Engineering, pp.155-165
\bibitem{}
Phipps,G; \emph{"Comparing Observed Bug and Productivity Rates for Java and C++" } in Software—Practice \& Experience; vol.29, issue 4, pp.345-358, April 10, 1999 
\bibitem{}
Prechelt, L., \emph{"An empirical comparison of seven programming languages"} in Computer , vol.33, no.10, pp.23-29, Oct 2000
\bibitem{}
Fourment,M.; Gillings,M.; \emph{"A comparison of common programming languages used in bioinformatics"} in BMC Bioinformatics Feb 2008
\bibitem{}
Furia,C.; Nanz,S.; \emph{"A Comparative Study of Programming Languages in Rosetta Code"} in ICSE '15, Proceedings of the 37th International Conference on Software Engineering, vol.1, pp.778-788
\bibitem{}
Iowa State University of Science and Technology,  \emph{"Boa - Mining Ultra-Large-Scale Software Repositories"}, http://boa.cs.iastate.edu/

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2016-02-17 main.tex
Changed the reference to .bib file

@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-\bibitem{}
-Hattori, L.P.; Lanza, M., \emph{"On the nature of commits"} in Automated Software Engineering - Workshops, 2008.
+%\begin{thebibliography}{1}
+
+\bibliographystyle{abbrv}
+\bibliography{hotfixes}
+
+%\bibitem{}
+%Hattori, L.P.; Lanza, M., \emph{"On the nature of commits"} in Automated Software Engineering - Workshops, 2008.
 ASE Workshops 2008.
@@
 2008
-\bibitem{}
-Livshits,B.; Zimmerman,T.; \emph{"DynaMine: Finding Common Error Patterns by Mining Software Revision Histories"} in Proceedings of the 10th European software engineering conference held jointly with 13th ACM SIGSOFT international symposium on Foundations of software engineering, pp.296-305
-\bibitem{}
-Mockus, A.; Votta, L.G., \emph{"Identifying reasons for software changes using historic databases"} in Software Maintenance, 2000.
+%\bibitem{}
+%Livshits,B.; Zimmerman,T.; \emph{"DynaMine: Finding Common Error Patterns by Mining Software Revision Histories"} in Proceedings of the 10th European software engineering conference held jointly with 13th ACM SIGSOFT international symposium on Foundations of software engineering, pp.296-305
+%\bibitem{}
+%Mockus, A.; Votta, L.G., \emph{"Identifying reasons for software changes using historic databases"} in Software Maintenance, 2000.
 Proceedings.
 International Conference on , vol., no., pp.120-130, 2000
-\bibitem{}
-Ratzinger,J.; Sigmund,T.; Gall,H.; \emph{"On the Relation of Refactoring and Software Defects"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.
+%\bibitem{}
+%Ratzinger,J.; Sigmund,T.; Gall,H.; \emph{"On the Relation of Refactoring and Software Defects"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.
 35-38
-\bibitem{}
-Hindle,A.; German,D.; Holt,R.; \emph{"What do large commits tell us?: a taxonomical study of large commits"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.99-108
-\bibitem{}
-Oliva,G.; Steinmacher,I.; Wiese,I.; Aurelio,M.; \emph{"What Can Commit Metadata Tell Us About Design Degradation?"} in IWPSE 2013, Proceedings of the 2013 International Workshop on Principles of Software Evolution, pp.18-27
-\bibitem{}
-Phillips,S.; Sillito,J.; Walker,R.; \emph{"Branching and merging: an investigation into current version control practices"} in CHASE '11 Proceedings of the 4th International Workshop on Cooperative and Human Aspects of Software Engineering, pp.9-15
-\bibitem{}
-Agrawal, K.; Amreen, S.; Mockus, A., \emph{"Commit Quality in Five High Performance Computing Projects"} in Software Engineering for High Performance Computing in Science (SE4HPCS), 2015 IEEE/ACM 1st International Workshop on , vol., no., pp.24-29, 18-18 May 2015
-\bibitem{}
-Steff, M.; Russo, B., \emph{"Co-evolution of logical couplings and commits for defect estimation"} in Mining Software Repositories (MSR), 2012 9th IEEE Working Conference on , vol., no., pp.213-216, 2-3 June 2012
-\bibitem{}
-Williams,C.; Spacco,J.; \emph{"Branching and Merging in the Repository"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.19-22
-\bibitem{}
-Kolassa,C.; Riehle,D.; Salim,M.; \emph{"The Empirical Commit Frequency Distribution of Open Source Projects"} in WikiSym '13, Proceedings of the 9th International Symposium on Open Collaboration, Article No.18
-\bibitem{}
-Ray,B.; Posnett,D.; Filkov,V.; Devanbu,P;  \emph{A Large Scale Study of Programming Languages and Code Quality"} in Github;  FSE 2014, Proceedings of the 22nd ACM SIGSOFT International Symposium on Foundations of Software Engineering, pp.155-165
-\bibitem{}
-Phipps,G; \emph{"Comparing Observed Bug and Productivity Rates for Java and C++" } in Software—Practice \& Experience; vol.29, issue 4, pp.345-358, April 10, 1999 
-\bibitem{}
-Prechelt, L., \emph{"An empirical comparison of seven programming languages"} in Computer , vol.33, no.10, pp.23-29, Oct 2000
-\bibitem{}
-Fourment,M.; Gillings,M.; \emph{"A comparison of common programming languages used in bioinformatics"} in BMC Bioinformatics Feb 2008
-\bibitem{}
-Furia,C.; Nanz,S.; \emph{"A Comparative Study of Programming Languages in Rosetta Code"} in ICSE '15, Proceedings of the 37th International Conference on Software Engineering, vol.1, pp.778-788
-\bibitem{}
-Iowa State University of Science and Technology,  \emph{"Boa - Mining Ultra-Large-Scale Software Repositories"}, http://boa.cs.iastate.edu/
+%\bibitem{}
+%Hindle,A.; German,D.; Holt,R.; \emph{"What do large commits tell us?: a taxonomical study of large commits"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.99-108
+%\bibitem{}
+%Oliva,G.; Steinmacher,I.; Wiese,I.; Aurelio,M.; \emph{"What Can Commit Metadata Tell Us About Design Degradation?"} in IWPSE 2013, Proceedings of the 2013 International Workshop on Principles of Software Evolution, pp.18-27
+%\bibitem{}
+%Phillips,S.; Sillito,J.; Walker,R.; \emph{"Branching and merging: an investigation into current version control practices"} in CHASE '11 Proceedings of the 4th International Workshop on Cooperative and Human Aspects of Software Engineering, pp.9-15
+%\bibitem{}
+%Agrawal, K.; Amreen, S.; Mockus, A., \emph{"Commit Quality in Five High Performance Computing Projects"} in Software Engineering for High Performance Computing in Science (SE4HPCS), 2015 IEEE/ACM 1st International Workshop on , vol., no., pp.24-29, 18-18 May 2015
+%\bibitem{}
+%Steff, M.; Russo, B., \emph{"Co-evolution of logical couplings and commits for defect estimation"} in Mining Software Repositories (MSR), 2012 9th IEEE Working Conference on , vol., no., pp.213-216, 2-3 June 2012
+%\bibitem{}
+%Williams,C.; Spacco,J.; \emph{"Branching and Merging in the Repository"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.19-22
+%\bibitem{}
+%Kolassa,C.; Riehle,D.; Salim,M.; \emph{"The Empirical Commit Frequency Distribution of Open Source Projects"} in WikiSym '13, Proceedings of the 9th International Symposium on Open Collaboration, Article No.18
+%\bibitem{}
+%Ray,B.; Posnett,D.; Filkov,V.; Devanbu,P;  \emph{A Large Scale Study of Programming Languages and Code Quality"} in Github;  FSE 2014, Proceedings of the 22nd ACM SIGSOFT International Symposium on Foundations of Software Engineering, pp.155-165
+%\bibitem{}
+%Phipps,G; \emph{"Comparing Observed Bug and Productivity Rates for Java and C++" } in Software—Practice \& Experience; vol.29, issue 4, pp.345-358, April 10, 1999 
+%\bibitem{}
+%Prechelt, L., \emph{"An empirical comparison of seven programming languages"} in Computer , vol.33, no.10, pp.23-29, Oct 2000
+%\bibitem{}
+%Fourment,M.; Gillings,M.; \emph{"A comparison of common programming languages used in bioinformatics"} in BMC Bioinformatics Feb 2008
+%\bibitem{}
+%Furia,C.; Nanz,S.; \emph{"A Comparative Study of Programming Languages in Rosetta Code"} in ICSE '15, Proceedings of the 37th International Conference on Software Engineering, vol.1, pp.778-788
+%\bibitem{}
+%Iowa State University of Science and Technology,  \emph{"Boa - Mining Ultra-Large-Scale Software Repositories"}, http://boa.cs.iastate.edu/
 
-\end{thebibliography}
+%\end{thebibliography}
 

### 2016-02-17 main.tex
Changed the style of the paper to sig-alternate (ICSE's format) and fixed all the references

@@
 
-\documentclass[conference]{IEEEtran}
+\documentclass{sig-alternate}
 \usepackage{blindtext, graphicx}
 \usepackage{fancyhdr} 
+\usepackage{txfonts}
 \graphicspath{ {images/} }
-\fancyhf{}
-\cfoot{\thepage}
-\pagestyle{fancy}
+
+\newcommand\FIXME[1]{\textbf{FIXME: #1}}
+%\fancyhf{}
+%\cfoot{\thepage}
+%\pagestyle{fancy}
 
@@
 %
-\ifCLASSINFOpdf
+%\ifCLASSINFOpdf
   % \usepackage[pdftex]{graphicx}
@@
   % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
+%\else
   % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
@@
   % \DeclareGraphicsExtensions{.eps}
-\fi
+%\fi
 
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+%\hyphenation{op-tical net-works semi-conduc-tor}
 
-\pagenumbering{arabic}
+%\pagenumbering{arabic}
 \begin{document}
+\conferenceinfo{MSR}{'16 Austin, Texas USA}
 %
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Investigating Fault-Proneness Through Hotfix Commits: A Research in the Light of Different Programming Languages}
+
+\title{Investigating Fault-Proneness Through Hotfix Commits:\\A Research in the Light of Different Programming Languages}
+\numberofauthors{4}
+\author{
+	Martin Lazarov, Ismaa'eel Ataullah, Vladyslav Kolesnyk, Veselin Pavlov\\
+	\affaddr{University College London}\\
+	\affaddr{London, UK}
+}
+
+%\title{Investigating Fault-Proneness Through Hotfix Commits: A Research in the Light of Different Programming Languages}
 
@@
 % affiliations
-\author{\IEEEauthorblockN{Martin Lazarov\IEEEauthorrefmark{1},
-Ismaa'eel Ataullah\IEEEauthorrefmark{2}, Vladyslav Kolesnyk\IEEEauthorrefmark{3},
-Veselin Pavlov\IEEEauthorrefmark{4}}
-\IEEEauthorblockA{Department of Computer Science,
-UCL\\
-Email: \IEEEauthorrefmark{1}martin.lazarov.12@ucl.ac.uk,
-\IEEEauthorrefmark{2}ismaaeel.ataullah.12@ucl.ac.uk,
-\IEEEauthorrefmark{3}vladyslav.kolesnyk.12@ucl.ac.uk,
-\IEEEauthorrefmark{4}veselin.pavlov.15@ucl.ac.uk}}
+%\author{\IEEEauthorblockN{Martin Lazarov\IEEEauthorrefmark{1},
+%Ismaa'eel Ataullah\IEEEauthorrefmark{2}, Vladyslav Kolesnyk\IEEEauthorrefmark{3},
+%Veselin Pavlov\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{Department of Computer Science,
+%UCL\\
+%Email: \IEEEauthorrefmark{1}martin.lazarov.12@ucl.ac.uk,
+%\IEEEauthorrefmark{2}ismaaeel.ataullah.12@ucl.ac.uk,
+%\IEEEauthorrefmark{3}vladyslav.kolesnyk.12@ucl.ac.uk,
+%\IEEEauthorrefmark{4}veselin.pavlov.15@ucl.ac.uk}}
 
@@
 In this paper, we propose analysing hotfix commits in various development areas as a means of indicating what languages tend to produce the least amount of problems after software has been published.
-We have focused on parsing metadata from projects hosted on SourceForge related to five specific topics (“software development”, “databases”, “security”, “games”, “front-ends”) and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages.
+We have focused on parsing metadata from projects hosted on SourceForge related to five specific topics (``software development'', ``databases'', ``security'', ``games'', ``front-ends'') and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages.
 \end{abstract}
 
-\begin{IEEEkeywords}
-Version Control, hotfixes, SourceForge, commits metadata.
-\end{IEEEkeywords}
+% A category with the (minimum) three required fields
+\category{D.2.7}{Software Engineering}{\FIXME{}}
+%A category including the fourth, optional field follows...
+
+%\terms{Experimentation, Measurement, Languages, Performance}
+
+\keywords{Version Control, hotfixes, SourceForge, commits metadata}
+
+%\begin{IEEEkeywords}
+%Version Control, hotfixes, SourceForge, commits metadata.
+%\end{IEEEkeywords}
 
@@
 Our research attempts to provide this crucial information by correlating hotfixes with the choice of language to develop software in a specific area.
-As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: “a corrective commit deployed within 1 day, following a project’s version update to fix critical issues in the release”.\par
+As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: ``a corrective commit deployed within 1 day, following a project's version update to fix critical issues in the release''.\par
 We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used.
@@
 We have also decided to follow this approach due to its openness and accessibility.
-This method is also very widely adopted amongst the MSR community [1].
+This method is also very widely adopted amongst the MSR community \cite{Hattori2008}.
 Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages’ stability.
@@
 A large number of the reports that we studied implemented a very simple approach.
-They focused on the commit message as the main mechanism to try to identify whether a commit was of a corrective type [1, 4, 5].
-They simply investigated if the string of text contained particular key words, such as: “bug”, “fix”, “issue”, “error”, “broke”, etc.\par
+They focused on the commit message as the main mechanism to try to identify whether a commit was of a corrective type~\cite{Hattori2008, Ratzinger2008, Hindle2008}.
+They simply investigated if the string of text contained particular key words, such as: ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.\par
 
@@
 Their algorithm consisted of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g.
-“for”, “code”, “to”, “the”, etc.
-) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits [3].\par
+``for'', ``code'', ``to'', ``the'', etc.
+) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits \cite{Mockus2000}.\par
 
-Another approach to identify bug fix commits involved performing a classification based on the size of the commits [1, 3, 5].
+Another approach to identify bug fix commits involved performing a classification based on the size of the commits \cite{Hattori2008,Mockus2000,Hindle2008}.
 The size of a commit can be measured based on the number of lines or the amount of files that were changed.
 According to these authors, large commits tend to improve code quality or add new features, while small ones are considered to be corrective - targeted at fixing bugs in the project.
-Some researchers proposed a further fine-grained size classification depending on the number of files changed per commit [1].
+Some researchers proposed a further fine-grained size classification depending on the number of files changed per commit \cite{Hattori2008}.
 They defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large (126 up ).
@@
 According to some authors, issues in critical components (e.g.
-errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered [2].
-In [3], the researchers argue that the time between changes will depend on the type of alteration.
+errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered \cite{Livshits2005}.
+In \cite{Mockus2000}, the researchers argue that the time between changes will depend on the type of alteration.
 Furthermore, they hold the view that corrective changes require the least amount of time to be completed as often they have to be performed promptly regardless of the complexity of the task.\par
 
-In terms of the time aspect, there was another paper which investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk [10].
+In terms of the time aspect, there was another paper which investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk \cite{Williams2008}.
 However, they explored branches in general without focusing specifically on bugfix branches.
 Some other scientists were trying to explain people’s behaviour during commits by investigating their commit frequency distribution.
-As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits [11].\par
+As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits \cite{Kolassa2013}.\par
 
-Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed [6].
+Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed \cite{Oliva2013}.
 However, they did not make their investigation language-specific as opposed to what we propose in our paper.\par
@@
 Other scientists tried to explore the logical coupling of commits as a way of identifying potential software defects.
-They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future [9].\par
+They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future \cite{Steff2012}.\par
 
@@
 According to the study performed by Phillips et al.
-[7], 41\% of the tested projects had incorporated an isolated bug fix branch.
+\cite{Phillips2011}, 41\% of the tested projects had incorporated an isolated bug fix branch.
 
 \subsection{Comparison of programming languages' stability}
-Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality [8].
+Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality \cite{Agrawal2015}.
 Furthermore, there were several papers that compared and contrasted the performance of different languages, which led to our second research topic.
@@
 
-There were researchers that focused on the effect of programming languages on software quality [12].
+There were researchers that focused on the effect of programming languages on software quality \cite{Ray2014}.
 The authors based performance on the rate of defect occurrences in various languages.
@@
 
-In [13] Java and C++ were evaluated by comparing bug rates.
+In \cite{Phipps1999}, Java and C++ were evaluated by comparing bug rates.
 In this case, there were two definitions for bugs - a defect and a bug itself.
@@
 
-There were researchers that used slightly different parameters to compare programming languages, these included program length, programming effort, run-time efficiency, memory consumption, and reliability [14].\par
+There were researchers that used slightly different parameters to compare programming languages, these included program length, programming effort, run-time efficiency, memory consumption, and reliability \cite{Prechelt2000}.\par
 
-An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages [15].
+An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages \cite{Nanz2015}.
 This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.\par
 
-Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness [16].
+Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness \cite{Dyer-Nguyen-Rajan-Nguyen-13} \FIXME{IS IT BOA's PAPER HERE?}.
 They argued that conciseness is an important factor as a more compact language can help write a program with fewer bugs.
@@
 \section{Solution}
-Our hypothesis was motivated by research paper [12]. In it the authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.\par
+Our hypothesis was motivated by research paper \cite{Ray2014}. In it the authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.\par
 
@@
 \par
-In order to evaluate our hypothesis, we developed a tool in Boa [17] that would allow us to incorporate our ideas and perform an analysis to verify them.
+In order to evaluate our hypothesis, we developed a tool in Boa \cite{Dyer-Nguyen-Rajan-Nguyen-13} that would allow us to incorporate our ideas and perform an analysis to verify them.
 Next, we present our reasoning and what factors we considered to accomplish our research.\par
@@
 Unfortunately, no prior projects implemented a technique to detect version bump commits.
-However, from the observed data it turned out to be quite straightforward to extract this information by generating a regular expression to match strings like: “bump”, “version”, “update” in an appropriate order.\par
+However, from the observed data it turned out to be quite straightforward to extract this information by generating a regular expression to match strings like: ``bump'', ``version'', ``update'' in an appropriate order.\par
 
-From our Related Work, we found projects [1, 4, 5], which distinguished bug fix commits only based on the commit messages.
+From our Related Work, we found projects \cite{Hattori2008,Ratzinger2008,Hindle2008}, which distinguished bug fix commits only based on the commit messages.
 However, due to the similarity between commit messages for normal bug fixes and hotfix commits, and occasionally the large sequences of text, it was challenging to use the same approach.
@@
 As previously discussed, we used a regular expression which examined commit messages for words associated with a corrective commit (e.g.
-“bug”, “fix”, “issue”, “error”, “broke”, etc.) to detect all bug fix commits.
+``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.) to detect all bug fix commits.
 Next, we considered other possibilities that would let us detect hotfixes in a more reliable way.
-Some of the papers we discussed in Related Work [1, 3, 5, 11] incorporated metadata, such as: the size of changes (the number of files that were altered) and the time of the commit (how long it took for a patch to be submitted) to identify bug fixes.
+Some of the papers we discussed in Related Work \cite{Hattori2008,Mockus2000,Hindle2008,Kolassa2013} incorporated metadata, such as: the size of changes (the number of files that were altered) and the time of the commit (how long it took for a patch to be submitted) to identify bug fixes.
 We considered these options for our purposes as well.
@@
 We decided to analyse SourceForge 2013 because it provided categorisation of projects into programming topics (e.g.
-code related to: “databases”, “software development”, “security”, etc.).
+code related to: ``databases'', ``software development'', ``security'', etc.).
 Unfortunately, Github did not possess such classification.
@@
 Therefore, we may have introduced some false positives or failed to account for actual hotfixes.
-  \item the way we detected bump commits was by writing a regular expression to match words like:  “bump”, “version”, “update” in an appropriate order.
+  \item the way we detected bump commits was by writing a regular expression to match words like:  ``bump'', ``version'', ``update'' in an appropriate order.
 Thus, we may have excluded version updates had they been indicated differently.
@@
 We expect that the results we obtained will help developers and project managers improve their planning and understand what the most appropriate languages are for their specific topic.
+
+\bibliographystyle{abbrv}
+\bibliography{hotfixes}
 
@@
 % by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
+%\ifCLASSOPTIONcaptionsoff
+%  \newpage
+%\fi
 
@@
 %\begin{thebibliography}{1}
-
-\bibliographystyle{abbrv}
-\bibliography{hotfixes}
 

### 2016-02-17 msr_paper/main.tex
Moved files used for MSR paper into the msr_paper folder

--- 2016-02-17 main.tex
@@
 This method is also very widely adopted amongst the MSR community \cite{Hattori2008}.
-Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages’ stability.
+Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages' stability.
 
@@
 The first major problem we had to solve was identifying actual hotfix commits.
-Due to Boa’s limitations, we were not able to find commit history from hotfix branch.
+Due to Boa's limitations, we were not able to find commit history from hotfix branch.
 We were only capable of collecting information with regard to commits performed on the main branch.
@@
 However, they explored branches in general without focusing specifically on bugfix branches.
-Some other scientists were trying to explain people’s behaviour during commits by investigating their commit frequency distribution.
+Some other scientists were trying to explain people's behaviour during commits by investigating their commit frequency distribution.
 As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits \cite{Kolassa2013}.\par
@@
 \centering
-\includegraphics[width=0.75\textwidth]{Bar_chart}
+\includegraphics[width=0.5\textwidth]{Bar_chart}
 \caption{Overall hotfixes per programming languages for 100 releases.}
@@
 Thus, we can say this aspect of our hypothesis is somewhat supported.
-It’s only partially justified because while Python, C and C++ have similar hotfix rates, Python still ranks at the bottom of our results table.\\
+It's only partially justified because while Python, C and C++ have similar hotfix rates, Python still ranks at the bottom of our results table.\\
 
@@
 To increase the significance of our study we performed both an overall estimation of hotfixes and a topic-specific analysis.
-The tool that we created was written in Boa and mined repositories from SourceForge 2013’s dataset.
-Our goal was to verify whether the general relationships between languages’ reliability holds with regard to hotfixes.
+The tool that we created was written in Boa and mined repositories from SourceForge 2013's dataset.
+Our goal was to verify whether the general relationships between languages' reliability holds with regard to hotfixes.
 Specifically, we investigated the following relations: strongly-typed vs weakly-typed languages, static vs dynamic and functional vs procedural.

### 2016-02-19 msr_paper/main.tex
First draft of paper

@@
 \usepackage{txfonts}
+\usepackage{lipsum}
+\usepackage{booktabs}
+
 \graphicspath{ {images/} }
@@
 Planning and resource allocation are arguably the most crucial components of any software project.
-Our research was motivated by the fact that very often, in the planning phase project managers neglect the importance of hotfixes – which becomes essential once a project is deployed.
-The most common reason for that practice is the fact that hotfixes are not constant and their presence is difficult for prediction.
-As a result, it is quite challenging to estimate their quantity and incorporate them into the cost of the project.
-Our research attempts to provide this crucial information by correlating hotfixes with the choice of language to develop software in a specific area.
+Our research was motivated by the fact that very often, in the planning phase project managers neglect the importance of hotfixes - which becomes essential once a project is deployed.
+
+Generally, a hotfix is related to correcting a critical problem which prevents a particular software product from performing as expected.
+This could mean a total or partial loss of service, a major security breach that has to be patched instantly, etc.
+As a result, such a fix is done outside of normal development and testing procedures.
+This characteristic of hotfixes implies that rapid development under urgent circumstances is usually required.
+Therefore, it can be often the case that lots of additional people and resources need to be allocated quickly for the correction, which leads to a significant increase in the cost of the overall product.
+
+Another aspect that makes hotfixes so important is their unpredictability which further prevents teams from planning in advance for such corrections into their development cycle.
+Moreover, hotfixes are sometimes delayed on purpose because people are afraid of the risk of applying the fix.
+All of these factors contribute to increasing the uncertainty and stress in managers and that can prevent them from taking the most appropriate action when a problem in production has to be resolved promptly.
+
 As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: ``a corrective commit deployed within 1 day, following a project's version update to fix critical issues in the release''.\par
+
+With our research we aim to help project managers and developers make a more informed decision when deciding what languages to use by comparing the number of hotfixes per programming language.
+To provide a better picture we also look at hotfixes per specific software topics.
+
 We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used.
@@
   \item a quantitative analysis identifying fault-proneness of specific programming topics and the relevant distribution of faults across programming languages.
-  \item an investigation into the following relationships between languages: strongly-typed vs weakly-typed, static vs dynamic and functional vs procedural to try to determine what group of languages is the most reliable.
+  \item an evaluation of the most used languages used across the top 16 most popular domains identified by SourceForge 13'.
+  \item an investigation into the languages that achieve the highest/lowest ratios of hotfix per version update.
+  \item research into the programming topics that experience the highest/lowest numbers of hotfix per version increase.
+  \item an investigation into the following relationships between languages: strongly-typed vs weakly-typed, static vs dynamic and scripting vs procedural to try to determine what group of languages is the most reliable.
   \item a foundation for future research in the area of identifying language-specific hotfixes.
@@
 
-An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages \cite{Nanz2015}.
+An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages \cite{Fourment2008}.
 This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.\par
 
-Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness \cite{Dyer-Nguyen-Rajan-Nguyen-13} \FIXME{IS IT BOA's PAPER HERE?}.
+Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness \cite{Nanz2015}.
 They argued that conciseness is an important factor as a more compact language can help write a program with fewer bugs.
@@
 \section{Solution}
-Our hypothesis was motivated by research paper \cite{Ray2014}. In it the authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.\par
 
-We were interested in investigating whether their findings applied when comparing number of hotfixes to total number of commits and if those conclusions were still relevant when we divided our project sources into various topics.\par
+% \begin{table}
+% \centering
+% \resizebox{\columnwidth}{!}{%
+% \begin{tabular}{r|lll}
+% \multicolumn{1}{r}{}
+% & \multicolumn{1}{l}{Heading 1}
+% & \multicolumn{1}{l}{Heading 2}
+% & \multicolumn{1}{l}{Heading 3} \\ \cline{2-4}
+% Row 1 & Cell 1,1 & Cell 1,2 & Cell 1,3 \\
+% Row 2 & Cell 2,1 & Cell 2,2 & Cell 2,3
+% \end{tabular}%
+% }
+% \end{table}
 
-More specifically we wanted to see whether:
-\begin{enumerate}
-\item \textbf{Strongly-typed languages required fewer hotfixes than weakly-typed languages.}
-\item \textbf{Statically-typed languages required fewer hotfixes than dynamically-typed languages.}
-\item \textbf{Functional languages required fewer hotfixes than procedural ones.}
-\end{enumerate}
-\par
-In order to evaluate our hypothesis, we developed a tool in Boa \cite{Dyer-Nguyen-Rajan-Nguyen-13} that would allow us to incorporate our ideas and perform an analysis to verify them.
-Next, we present our reasoning and what factors we considered to accomplish our research.\par
+\begin{table*}[!htbp]
+  \centering
+  \begin{tabular}{cccccccccccccccccc}
+    Language &
+    \rotatebox{90}{Databases} & \rotatebox{90}{Systems Administration} & 
+    \rotatebox{90}{Dynamic Content} & 
+    \rotatebox{90}{Front Ends} & 
+    \rotatebox{90}{Site Management} & 
+    \rotatebox{90}{Role Playing} & \rotatebox{90}{Games} & \rotatebox{90}{Security} &
+    \rotatebox{90}{Internet} & \rotatebox{90}{Software Development} & 
+    \rotatebox{90}{Testing} & 
+    \rotatebox{90}{Build Tools} & 
+    \rotatebox{90}{Code Generators}&
+    \rotatebox{90}{Communications} & \rotatebox{90}{Education} & \rotatebox{90}{Frameworks} &
+    \rotatebox{90}{Total Average} \\
+    \midrule
+    Java & 0.25 & 0.15 & 0.22 & 0.08 & 0.28 & 0.32 & 0.07 & 0.22 & 0.38 & 0.17 & 0.23 & 0.10 & 0.06 & 0.09 & 0.10 & 0.09 & \textbf{0.176} \\
+    Javascript & 0.26 & - & 0.27 & - & 0.17 & - & 0.17 & 0.23 & - & - & - & - & - & - & 0.18 & 0.18 & \textbf{0.209} \\
+    Perl & 0.17 & 0.05 & 0.22 & 0.44 & - & 0.41 & - & 0.06 & 0.13&0.24&-&-&-&0.17&0.19&0.32 & \textbf{0.218} \\
+    PHP & 0.27&0.23&0.27&0.28&0.24&-&-&0.23&0.10&0.38&-&-&0.14&-&0.67&0.31&\textbf{0.282} \\
+    C++ &0.24&0.12&0.26&0.06&-&0.51&0.35&0.31&0.20&0.26&0.27&0.35&0.73&0.08&0.12&0.43 & \textbf{0.286} \\ 
+    C &0.29&0.24&0.29&0.33&0.43&0.56&0.39&0.05&0.03&0.34&0.27&0.33&0.73&0.07&0.23& - & \textbf{0.305}\\
+    C\# & -&0.20&-&-&0.30&0.89&0.14&0.27&0.60& 0.08 & - & 0.07 & - & 0.57 & - & 0.27 & \textbf{0.339} \\
+    \bottomrule
+    Total Average & 0.25&0.17&0.28&0.28&0.28&0.47&0.33&0.17&0.12&0.32&0.27&0.34&0.53&0.11&0.27&0.32
+  \end{tabular}
+  \caption{Blabla}
+  \label{tab:1}
+\end{table*}
 
-To calculate the frequency of hotfixes for various projects and topics, we had to design a way of distinguishing a hotfix based on the commit messages.
-Based on the definition of a hotfix, we focused on discovering corrective commits that were submitted quickly after a new version release.
-In this way, we were hoping to identify problems specifically related to the version update.
-Therefore, we decided to divide our implementation into: detecting a \textbf{version bump commit} (i.e.
-a commit indicating a version increase) and identifying subsequent \textbf{bug fix commits}.\par
+Our work was inspired by the previous research paper \cite{Ray2014}, where authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.
+Our contribution primarily involved investigation of the fault-proneness of different languages based on the newly defined criteria - hotfixes per version update.
+Our dataset consist from [totalnumberofprojects] projects, selected from the top 19 most popular project types in the SourceForge dataset.
+That allows us to differentiate results for the different kinds of projects.\par
+Specifically, our investigation aimed to evaluate the following research questions:
 
-Unfortunately, no prior projects implemented a technique to detect version bump commits.
-However, from the observed data it turned out to be quite straightforward to extract this information by generating a regular expression to match strings like: ``bump'', ``version'', ``update'' in an appropriate order.\par
+\textbf{1) What languages are the most widely used across the top 16 most popular categories?} (as identified by SourceForge13 Dataset)
 
-From our Related Work, we found projects \cite{Hattori2008,Ratzinger2008,Hindle2008}, which distinguished bug fix commits only based on the commit messages.
-However, due to the similarity between commit messages for normal bug fixes and hotfix commits, and occasionally the large sequences of text, it was challenging to use the same approach.
-Therefore, in addition to bug fix commit detection, we applied some other parameters to ensure that it is not just a regular fix, but a hotfix.
-As previously discussed, we used a regular expression which examined commit messages for words associated with a corrective commit (e.g.
-``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.) to detect all bug fix commits.
-Next, we considered other possibilities that would let us detect hotfixes in a more reliable way.
-Some of the papers we discussed in Related Work \cite{Hattori2008,Mockus2000,Hindle2008,Kolassa2013} incorporated metadata, such as: the size of changes (the number of files that were altered) and the time of the commit (how long it took for a patch to be submitted) to identify bug fixes.
-We considered these options for our purposes as well.
-Consequently, in order to increase our confidence level that the current fix commit is an actual hotfix, we introduced the following parameters to verify the emergency of the fix:
-\begin{itemize}
-  \item \textbf{Time} - commits have to be submitted within 1 day from the version bump commit
-  \item \textbf{Look-ahead} range for commits - only n numbers of commits should be considered after version bump
-\end{itemize}
+Research on the most widely used languages across the top 16 most popular categories showed that there are surprisingly few such languages.
+We identified only 6 widely used languages: C\#, Perl, Python, C, C++ and PHP.
+The criteria of choosing the languages is based on whether the language was present in at least one half of the categories.
+One would expect Javascript to appear in the list.
+However, it is present only in 7 categories which is less than 50\% of the categories.
+This may be due to the fact that Javascript is mainly used for website development.
+The two most widely used languages from the list are C and C++ which are present in 15 out of 16 categories.
+While C++ is not present in site management, C is not present in frameworks.
+The rest of the languages in the list are present on average in 10 categories.
+Overall, the majority of these languages were not present in testing, code generators and site management categories.
+The observations made are reasonable since using these languages in such categories is very uncommon.
 
-\par
-The size of the commit (i.e.
-Number of files changed ) could have been used as another parameter to determine hotfixes as our investigation discovered that corrective type of changes generally affect few files (e.g.
-revert new alterations or switch off a problematic feature).
-However, based on our results, we found examples where lots of changes were merged into one commit.
-Consequently, several hotfixes may be submitted as part of a single merge commit, hence affecting many files.
-In such case, we would fail to account for these hotfix commits because of the large number of files involved in the fix.\par
+\textbf{2) What categories achieve the highest/lowest number of hotfixes per version update?}
 
-The reason why we decided to put a limit and investigate commits only within 1 day after a release was because of the nature of hotfixes.
-Whenever a live product contains a critical fault, it tends to be identified quickly as key functionalities stop working.
-Therefore, leaving bugs in production amplify the risk of the system being unusable.
-As a result, such hotfixes are assigned the highest priority and have to be performed rapidly even outside of the normal development process.\par
+In this section we investigate which categories achieve the highest or lowest numbers of hotfixes per version update.
+We proceed to analyse why certain categories might have low numbers of hotfixes and why others might require higher numbers of hotfixes.
 
-Furthermore, it is very likely that fixes performed on the first day after a new release will refer to problems occurring before the version increase (or at latest to commits submitted on that day).
-Therefore, we decided to focus on the Look-ahead to investigate the first n commits after a version update.
-We focused on the first 15 commits as anything after this point would increase the probability that a fix is related to code added after the release rather than before that.
-To provide more fine-grained results, we looked at the initial 5, 10 and 15 commits separately.\par
+If we look at the average number of hotfixes per category we see that the Internet, Security, communications and systems administration have the lowest average number of hot fixes per update out of the entire group of categories.
 
-With regard to evaluating the most optimal Look-ahead value, we decided to count the number of Forward Engineering commits as well (i.e.
-commits that introduce new features).
-We did this because we were interested in identifying the largest fix commits/forward commits ratio.
-This would allow us to guarantee that lower number of new features were introduced into the code and consequently that some of the higher number of fixes refer to the previous version.\par
+The categories with the highest average numbers of hotfixes per version update are Code Generation and Role playing which have 0.53 and 0.47 average hot fixes per update respectively.
+Both these values are much higher than the next highest (0.34) So we will only consider these two.
 
-Based on our research and the experiments we ran, we managed to identify the most optimal parameters (i.e.
-\textbf{time}: within 1 day after a release and \textbf{look-ahead}: the first 10 commits - \textit{see Table 1 in Appendix}).
-Having accomplished that, we focused on selecting a dataset.
-We decided to analyse SourceForge 2013 because it provided categorisation of projects into programming topics (e.g.
-code related to: ``databases'', ``software development'', ``security'', etc.).
-Unfortunately, Github did not possess such classification.
-We believed that being able to sort projects based on topics would produce greater insights and more useful information for developers, as they would be more interested in language efficiency for particular software development area.\par
+With regards to those with lower average hot fixes per update it is easy to see why.
+Applications under the categories as Security, Internet, Communications and systems administration are all information critical i.e.
+they have critical assets that need to be protected in order to maintain customer trust and keep the reputation of the creator intact.
+For example with communications applications it is vital that security is implemented in such a way that an adversary cannot intercept or obtain messages only intended for the receiver.
+This means that applications that fall under such categories should be extensively tested before deployment reducing the chance of a hot fix being needed after deployment.
 
-From all the available projects, we only concentrated on the ones that had incorporated versioning (i.e.
-containing bump commits).
-Our argument was that only if there were version updates which we could detect, we would be able to perform a hotfix analysis.
-Therefore, we extracted all the projects that had incorporated bump commits and applied our parameters to try to recognise the hotfixes.
-Consequently, for each programming language we analysed the number of hotfix and bump commits so that we would be able to
-\begin{enumerate}
-\item \textbf{Strongly-typed languages required fewer hotfixes than weakly-typed languages.}
-\item \textbf{Statically-typed languages required fewer hotfixes than dynamically-typed languages.}
-\item \textbf{Functional languages required fewer hotfixes than procedural ones.}
-\end{enumerate}
+If we look further into the results we can see that for example the usage of C in Internet and security categories has an incredibly low hot fix average (0.03 and 0.05 respectively).
+This could be down to fully comprehensive testing techniques, but could also be attributed to the fact that within those categories, the language has been used for an extensive period and hence there is a lot of experience with a variety of different bugs, meaning they are less likely to occur during development as experienced programmers are less likely to make those particular mistakes in the first place.
+This could also be the reason the projects written in Perl have a very low hot fix average for all four categories.
 
-\subsection{Overall performance of programming languages}
-We started with investigating the overall reliability of different languages regardless of the topic they belonged to.\\
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{Bar_chart}
-\caption{Overall hotfixes per programming languages for 100 releases.}
-\label{fig:my_label}
-\end{figure}
+When considering why role playing applications have a high average number of hot fixes we need to try and speculate as to why the hot fix rate was so much higher than that of projects that fall under the games category.
+It is possible that the games category is a general category are a lot simpler than those that fall under the role playing category which is why bugs might be more prominent as a result of increased complexity of the game.
+If we look a little closer at the results for the role playing category it becomes clear that for all languages there is a high average number of hot fixes per update which supports our theory.
 
-\par
-By performing a general analysis on all languages regardless of their topic, we compared the ratios of hotfix commit/release based on the available data.
-According to our results, Scala is the most reliable programming language with only 6.8 hotfixes per 100 releases ( \textit{see Table 1 in Appendix for more details}).
-On the other hand, Matlab is the worst performing language with 37.93 hot corrections per 100 version updates.\\
+If we look at Code generators on the other hand we see that for two languages (C and C++) the hot fix average is very high and for two other languages (PHP and Java) it is very low.
+It is not easy to conclude why this might be the case, however one can speculate that PHP and Java are established languages for writing code generators and due to this engender low hot fix averages.
+C and C++ might share syntactic details and semantic structure that would result in a similar hot fix average between them.
 
-\textbf{Hypothesis 1}:\par
-The strongly-typed languages we considered were: C\#, Java, Python, Scala.\par
-The weakly-typed languages we identified were: C, C++, Objective C, Perl, Javascript, PHP.\par
+\textbf{3) What languages achieve the highest/lowest number of hotfixes per version update?}
 
-First, Scala performs better than all weakly-typed languages.
-However, C\#, Java and Python do not support the hypothesis and do not outperform all weakly-typed languages.
-Moreover, all of them have a higher hotfix per version bump ratio than Objective C.
-Therefore, based on our results, we can state that it is often the case that a strongly-typed language \textbf{does not} have fewer hotfixes than a weakly-typed language and our analysis is inconclusive on this matter.\\
+Language comparison based on the average hotfix ratio over all topics has shown that Java has appeared to have the least number of hotfixes per version update (0.18).
+Javascript (0.21) and Perl (0.22) were found to perform slightly worse with the hotfix ratio value being just around 3-4\% higher.
+However, all the mentioned languages showed marginally better results with 8\% difference comparing to other common languages taken under analysis - PHP(0.28), C++(0.29), Python(0.30), C(0.31) and C\#(0.34).
 
-\textbf{Hypothesis 2}:\par
-Static languages considered: C, C++, Objective C, Java, Scala.\par
-Dynamic languages identified: Python, Perl, PHP, Javascript.\par
+\textbf{4) Are there any patterns based on the type of the programming language?}
 
-C performs better than all dynamic languages apart from Perl. 
-C++ is only more reliable than PHP. In all other cases, C++ contains more hotfixes than the other dynamically-typed languages. 
-Objective C is a very strong static candidate and is more reliable than all dynamic languages.
-Java performs slightly worse, but it still has overall dominance over the dynamically-typed languages.
-It has more amount of hotfixes per release only when compared with Perl.
-However, in the other cases, it is more efficient than all languages from the second group.
-Scala, as shown already is the best performer throughout all languages.
-Therefore, it is more reliable than all dynamic ones.
-From this analysis, we could say that in most of the cases, the second part of our hypothesis \textbf{is supported} by the results we obtained.\\
+The fourth research question that we decided to investigate was related to comparing the type of each programming language.
+The classification was inspired by Ray et al.
+[14], where languages were compared based on the number of bug fixes in each language and were divided into three main Classes.
+The first class was named ‘Programming Paradigm’ and it split languages into the following categories: Procedural, Functional, Scripting.
+The second criteria was Compilation Class, which yielded these types: Static and Dynamic.
+The final one, Type Class, separated languages into: Strongly-typed and weakly-typed.
 
-\textbf{Hypothesis 3}:\par
-Unfortunately, from the languages that we were able to collect information about, only Scala is a member of the functional programming languages.
-As mentioned already Scala is the best performer amongst all evaluated languages, and as such will undoubtedly have less hotfixes than all the procedural languages (e.g.
-C, C++,C\#, Objective C, Java ).
-Thus, based purely on our results, the third part of our hypothesis agrees with the data that we collected.\par
+According to the results that were achieved in the study [14], strongly-typed languages experienced fewer bug fixes than weakly-typed.
+Static programming languages required fewer corrective commits than dynamic ones.
+And finally, functional languages encountered fewer bug fixes than procedural ones.
 
-Overall, our hypothesis agreed with the results we obtained apart from a small number of cases.
-However, we expect that this may be due to the different amount of releases and hotfixes identified per language.
-We could also conclude that Functional-Static-Strong languages are better than Procedural-Dynamic-Weak ones.
-Therefore, we could also argue that by using combinations of the former type of languages, developers will be able to guarantee the lowest amount of hotfixes per release.
-As a result, the effort needed for maintenance will be reduced and software quality will be improved.
+Due to the fact that we selected to evaluate languages that appeared in at least 50% of the selected categories, we discovered only 6 languages and none of them was functional.
+Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains.
+Consequently, we evaluated the relationship between Procedural and Scripting languages instead of Procedural against Functional.
 
-\subsection{Topic: Databases}
-We take into account reliability as we want to reach reasonable scientific conclusions based on our findings.
-Unfortunately, we experience limitations within the data.
-For example, there are languages where we have identified only 2 or 3 bump commits.
-Deriving results from such a small dataset is extremely unreliable and it is not likely we can reach accurate solutions with them, hence we exclude them as outliers.
-These languages are Zope and C\# 1.0.\par
-We can see from our results that C has the lowest number of hotfix to bump commit ratio, suggesting that C is the most reliable language to program in, closely followed by Python.
-C++, PHP and Javascript have similar reliability according to our results but are less stable when compared to Perl or Python.
-Given the large number of bump commits for C++, PHP and Javascript, we can assume these results are reliable.\par
-Our hypothesis had three aspects to it and we will analyse whether our results support all of them.\\
+Analysing the first language relationship - Procedural (Java, C, C\#, C++) vs Scripting (Perl, Python, PHP), we were not able to detect any patterns across the categories.
+The reason was that the number of languages was too low to make any firm conclusions.
+It was always the case that at least one of the languages would not follow a pattern or the comparison would consider only one language from a certain category.
 
-\textbf{Hypothesis 1}:\par
-The strongly typed languages we consider are: Java, Python.\par
-The weakly typed languages we consider are: C, C++, Perl, PHP, Javascript.\par
-In this case we have a mixed series of results which suggest that this part of our hypothesis may not be entirely correct.
-Although Python is ranked as the second most reliable language, it is topped by C, which is a weakly-typed language.
-If Java had been placed third,  the hypothesis would have been better supported as we could have then treated C as an outlier.
-However, this is not the case.
-Above Java is another weakly-typed language, namely Perl.
-In conclusion, when considering database-specific projects, we cannot say that Strongly-typed languages required fewer hotfixes when compared to weakly-typed ones.\\
+Evaluating the second relationship between Static (Java, C, C\#, C++) and Dynamic (Perl, Python, PHP) languages led to the same conclusions.
+The reason was because the languages in these two categories coincided with the languages in the first two classes (Procedural and Scripting).
 
-\textbf{Hypothesis 2}:\par
-The statically-typed languages we consider are: C, Java, C++.\par
-The dynamically-typed languages we consider are: Javascript, Python, Perl, PHP.\par
+The third aspect that we investigated referred to comparing strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.
+Similarly to the first language relationship, we encountered that there was no clear pattern across the categories.
+Furthemore, we found the same problem of having comparisons against against only one member of a language class.
+Solely in the Communications category, we were able to detect some pattern, where weakly typed languages performed better.
+However, there were only two languages labelled as weak.
+Therefore, we cannot reach any satisfying conclusions based on this finding.
 
-Again we have a mixed list of results.
-C is top and this would suggest that statically-typed languages indeed do require fewer hotfixes when compared to dynamically-typed languages.
-However, the other static languages, Java and C++, are in the bottom half of the table in terms of reliability.
-Once again, we cannot say that this part of our hypothesis is supported when comparing languages used for developing database-specific programmes.\\
-
-\textbf{Hypothesis 3}:\par
-We cannot compare this aspect of the hypothesis as no languages in this set of results are classified as Functional languages.\par
-\textbf{In conclusion},no elements of our hypotheses were supported by our results.
-In two cases, our observations were not definitive and in the other we did not have the required results to make a comparison.
-We can conclude that our hypothesis is not supported when considering languages used for developing  database-related projects.
-
-\subsection{Topic: Games}
-In order to only analyse those results we deem reliable, we will exclude C\#, Javascript and Objective C on the basis that the number of bump commits analysed for them were too low.\par
-
-We can see from our results that for programming games, Java seems to be the most reliable language to use.
-C is second; however, it is somewhat behind, with 24.6\% of all commits being hotfix commits.
-C and C++ are very similar and are only separated by 0.1\%.
-In last place is Python which has a hotfix rate of almost 30\% and according to our results is the least reliable language to use.\\
-
-\textbf{Hypothesis 1}:\par
-The strongly-typed languages we consider are: Java, Python.\par
-The weakly-typed languages we consider are: C, C++.\par
-
-Java is the most reliable language by far according to our results.
-C, C++ and Python have quite similar hotfix rates.
-Java is a strongly-typed language which has a definitive advantage over the others.
-Thus, we can say this aspect of our hypothesis is somewhat supported.
-It's only partially justified because while Python, C and C++ have similar hotfix rates, Python still ranks at the bottom of our results table.\\
-
-\textbf{Hypothesis 2}:\par
-
-The statically-typed languages we consider are: C, Java, C++.\par
-The dynamically-typed languages we consider are: Python.\par
-
-This aspect of the hypothesis is categorically supported by our results, all statically-typed languages are ranked higher than the dynamically-typed Python in our results.
-We can therefore conclude that static languages required fewer hotfixes when compared to dynamic languages and hence the latter are less reliable when programming games.\\
-
-\textbf{Hypothesis 3}:\par
-We cannot compare this aspect of the hypothesis as no languages in this set of results are classed as Functional languages.\par
-
-\textbf{In conclusion}, we can state that our overall hypothesis is somewhat supported with regard to projects labeled as Games.
-
-\subsection{Topic: Front-ends}
-Here we can see a huge difference in numbers of bump commits detected in various projects.
-Thus, we can argue that not many projects were observed to have a Front-end label for sufficient analysis.
-And many other popular Front-end languages (e.g.
-Javascript, C\#) were not identified.\par
-Hence, after removing languages with a small number of detected projects, we ended up with \textbf{Scala} and \textbf{Java}, where the former has a slightly smaller hotfixes per bump ratio.\\
-
-\textbf{Hypothesis 1}:\par
-Both Scala and Java are considered to be strongly-typed; thus, no comparison can be made within the filtered languages. 
-Regarding all the languages, we can observe that Java and Scala, overall, have more reliable statistics than weakly-typed (PHP, C, Perl), hence the hypothesis is \textbf{satisfied}.\\
-
-\textbf{Hypothesis 2}:\par
-Both Scala and Java are classified as statically-typed; therefore, no comparison can be made within the filtered languages.
-In terms of all the languages in the list, we can observe that most of the static ones (C++, Scala and Java) are more reliable than dynamic languages (PHP, Perl), hence the hypothesis is \textbf{satisfied}.\\
-
-\textbf{Hypothesis 3}:\par
-From the set of languages, only Scala is considered to be functional and it indeed performed better than all the others, therefore the hypothesis is \textbf{satisfied}.
-
-\subsection{Topic: Software Development}
-We filtered languages that had insignificant number of projects under the Software Development label and ended up analysing: Java, C++, Python and C languages.
-It turns out Java appeared to be the most stable language, where on average for 10 version updates, only 1 hotfix would be detected.\\
-
-\textbf{Hypothesis 1}:\par
-There was no enough evidence to suggest that \textbf{all} strongly-typed languages (Java, Python) performed better than the weakly-typed ones (C++, C) within the Software Development topic.
-For example, Python(22.22\%) has more hotfixes than C++ (19.05\%).
-However, the average ratio for strongly-typed languages (16.5\%) is lower than the average ratio for weakly-typed (21.63\%).
-Thus, we can conclude that the hypothesis is \textbf{satisfied} in general.\\
-
-\textbf{Hypothesis 2}:\par
-Not all static languages would satisfy the hypothesis, as for example, Python as a dynamic language outperforms C (defined as a statically-typed language).
-However, by considering the average ratio of filtered languages, we can see that static languages (Java, C++, C - 19.1\%) by far outweigh the dynamic language Python - 22.2\%.Hence, the hypothesis is \textbf{satisfied}, however \textbf{dynamic languages are not represented} well in the dataset.\\
-
-\textbf{Hypothesis 3}:\par
-No functional languages were presented in dataset. Therefore, we cannot provide any conclusive argument with regard to the third component of our hypothesis.
-
-\subsection{Topic: Security}
-Due to the low number of releases per language that we were able to identify, we cannot guarantee that the results we collected would hold in general.
-However, in order to be able to perform some analysis on the data and reach somewhat valid results, we decided to exclude only the outlier languages  (i.e.
-the ones containing the lowest number of version updates).
-Thus, we decided to eliminate C\# and Python from the list of evaluated languages.\par
-
-According to our filtered data, the best performing language is C with just 1 hotfix per 20 releases on average.
-On the other hand, the least reliable one is C++ as it requires a hotfix for approximately every 6.5 releases.\\
-
-\textbf{Hypothesis 1}:\par
-The analysis of our results reveals that strongly-typed languages (Java) do not perform better in comparison with weakly-typed languages (C, C++, Javascript, Perl, PHP).
-In fact, Java only outperforms C++, but in all other cases, it requires more hotfixes than the remaining weakly-typed languages (C, Javascript, Perl, PHP).
-Therefore, the first part of our hypothesis does not hold for projects related to Security.\\
-
-\textbf{Hypothesis 2}:\par
-Our observations were inconclusive on whether static languages (C, C++, Java) are more reliable than dynamic ones (Javascript, Perl, PHP).
-Based on our results, C performed better than all dynamically-typed languages.
-However, Java and C++ both required more hotfixes per version update than all dynamic languages.
-Therefore, our analysis cannot provide a convincing argument whether static languages are more reliable than dynamic ones.
-As a result, the second part of our hypothesis is inconclusive on this matter.\\
-
-\textbf{Hypothesis 3}:\par
-Unfortunately, from the results that we obtained, we could not find any functional languages to compare them to procedural ones (C, C++, Java).
-Consequently, regarding the languages used in Security related projects, we are not capable of deriving any conclusions about the required number of hotfixes whatsoever.
-Thus, there is no argument to support or disprove the third element of our hypothesis.\par
-Overall, due to the insufficient number of releases in our data, we were not able to state conclusively whether our hypothesis is supported or not.
-Although there were examples of disproving it, we believe that having information about a larger amount of releases may actually reveal results closer to the overall observations.
+Overall, we could not identify any convincing patterns based on the class comparison between languages mostly due to the low number of extracted programming languages.
 
@@
   \item most open source projects are not under the stress of releasing as opposed to commercial projects. Consequently, they do not release very often and/or do not support versioning.
+  \item a limited number of programming topics was used (16 in total), which did not allow us to observe distinguishable patterns.
+  \item a low amount of languages satisfied our predicate (being present in at least one half of all evaluated topics), which did not allow us to state any conclusive results.
   \item only one dataset was used to evaluate our hypothesis - SourceForge 2013.
@@
 \section{Conclusions and Future Work}
-We have presented a research in detecting the fault-proneness of programming languages through hotfix commits - a topic that has not been investigated so far.
-To increase the significance of our study we performed both an overall estimation of hotfixes and a topic-specific analysis.
+We have presented a study in detecting the fault-proneness of programming languages through hotfix commits - a topic that has not been investigated so far.
+As part of our research we performed a topic-specific analysis across 16 programming domains to detect the most popular languages used in the evaluated  projects.
+We investigated languages and topics that achieved highest/lowest ratios of hotfix per version update and tried to identify patterns based on the language classification.
 The tool that we created was written in Boa and mined repositories from SourceForge 2013's dataset.
 Our goal was to verify whether the general relationships between languages' reliability holds with regard to hotfixes.
-Specifically, we investigated the following relations: strongly-typed vs weakly-typed languages, static vs dynamic and functional vs procedural.
-We expected that in each pair the first group of languages will be dominant.
-Although our overall results supported our hypothesis, on a topic-by-topic basis we were not able to verify it.\par
-Therefore, we believe that some future work has to be accomplished to increase the validity of our observations.
+Our results identified 6 languages as being the most widely used across different domains - C\#, Perl, Python, C, C++ and PHP.
+We discovered that Java achieves the least number of hotfixes per version update (0.18), while C\# requiring most amount of hotfixes per version increase (0.34).
+From the different domains, we recognized: Security, Internet and Communications as being the most reliable (i.e.
+requiring least amount of hotfixes per version update) while Role Playing and Code Generation required highest number of corrections.
+Regarding the comparison between language classes, we were not able to identify any conclusive patterns.\par
+There are few aspects that could be improved as Future Work.
 In particular, more datasets could be used, including Github (provided that projects are categorised into topics).

### 2016-02-19 msr_paper/main.tex
fixed table caption
Updated the structure of the paper and worked on the results section

@@
 \usepackage{booktabs}
+\usepackage{gitinfo}
 
@@
 %\boldmath
-Choosing the most suitable tools and programming languages is a very challenging task for every project manager and developer.
+Choosing the most suitable programming language is a very challenging task for every project manager and developer.
 Often the success of a software project is determined by the stability of the system after a release.
 In this paper, we propose analysing hotfix commits in various development areas as a means of indicating what languages tend to produce the least amount of problems after software has been published.
-We have focused on parsing metadata from projects hosted on SourceForge related to five specific topics (``software development'', ``databases'', ``security'', ``games'', ``front-ends'') and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages.
+We have focused on parsing metadata from the 16 most popular projects hosted on SourceForge and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages.
 \end{abstract}
@@
 \end{itemize}
-\par
-\textbf{Structure of the paper.}\\
-We discuss Related Work in section 2.
-The next segment discusses our Solution.
-Following from that, an analysis of the gathered results is presented.
-Finally, we review our approach and findings, and examine Future Work.
+
+The rest of the paper is organised as follows: Related work is discussed in section 2.
+Section 3 explains our empirical study with 4 research questions.
+In Section 4, the results are presented and discussed and followed by our conclusion and future work in Section 5.
 
@@
 
-\section{Solution}
+\section{Empirical Study}
 
@@
 
+Our work was inspired by the previous research paper \cite{Ray2014}, where authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.
+Our contribution primarily involved investigation of the fault-proneness of different languages based on the newly defined criteria - hotfixes per version update.
+
+\subsection{Dataset}
+The SourceForge13 Dataset is selected for this study.
+It consists of \FIXME{[totalnumberofprojects]} projects, selected from the top 16 most popular project types in the SourceForge dataset.
+That enables us to differentiate results for the different kinds of projects.
+
+\subsection{Research Questions}
+\noindent Our study aimed to answer the following research questions:
+
+\noindent\textbf{RQ1 (Language Popularity):} What languages are the most widely used across the top 16 most popular categories?
+
+\noindent\textbf{RQ2 (Hotfixes vs Project Categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
+
+\noindent\textbf{RQ3 (Hotfixes vs Languages):} What languages achieve the highest/lowest number of hotfixes per version update?
+
+\noindent\textbf{RQ4 (Hotfixes vs Language Types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?
+
 \begin{table*}[!htbp]
 	\centering
-  \begin{tabular}{cccccccccccccccccc}
-    Language &
+	\small
+	\begin{tabular}{c c c c r r r r r r r r r r r r r r r r r}
+		Language & $\mathcal{P}$ & $\mathcal{C}$ & $\mathcal{T}$ & 
 		\rotatebox{90}{Databases} & \rotatebox{90}{Systems Administration} & 
@@
 		\rotatebox{90}{Total Average} \\
-    \midrule
-    Java & 0.25 & 0.15 & 0.22 & 0.08 & 0.28 & 0.32 & 0.07 & 0.22 & 0.38 & 0.17 & 0.23 & 0.10 & 0.06 & 0.09 & 0.10 & 0.09 & \textbf{0.176} \\
-    Javascript & 0.26 & - & 0.27 & - & 0.17 & - & 0.17 & 0.23 & - & - & - & - & - & - & 0.18 & 0.18 & \textbf{0.209} \\
-    Perl & 0.17 & 0.05 & 0.22 & 0.44 & - & 0.41 & - & 0.06 & 0.13&0.24&-&-&-&0.17&0.19&0.32 & \textbf{0.218} \\
-    PHP & 0.27&0.23&0.27&0.28&0.24&-&-&0.23&0.10&0.38&-&-&0.14&-&0.67&0.31&\textbf{0.282} \\
-    C++ &0.24&0.12&0.26&0.06&-&0.51&0.35&0.31&0.20&0.26&0.27&0.35&0.73&0.08&0.12&0.43 & \textbf{0.286} \\ 
-    C &0.29&0.24&0.29&0.33&0.43&0.56&0.39&0.05&0.03&0.34&0.27&0.33&0.73&0.07&0.23& - & \textbf{0.305}\\
-    C\# & -&0.20&-&-&0.30&0.89&0.14&0.27&0.60& 0.08 & - & 0.07 & - & 0.57 & - & 0.27 & \textbf{0.339} \\
-    \bottomrule
-    Total Average & 0.25&0.17&0.28&0.28&0.28&0.47&0.33&0.17&0.12&0.32&0.27&0.34&0.53&0.11&0.27&0.32
+		\hline
+		Java & P & S & S & 0.25 & 0.15 & 0.22 & 0.08 & 0.28 & 0.32 & 0.07 & 0.22 & 0.38 & 0.17 & 0.23 & 0.10 & 0.06 & 0.09 & 0.10 & 0.09 & \textbf{0.176} \\
+		Javascript  & & & & 0.26 & - & 0.27 & - & 0.17 & - & 0.17 & 0.23 & - & - & - & - & - & - & 0.18 & 0.18 & \textbf{0.209} \\
+		Perl  & & & & 0.17 & 0.05 & 0.22 & 0.44 & - & 0.41 & - & 0.06 & 0.13&0.24&-&-&-&0.17&0.19&0.32 & \textbf{0.218} \\
+		PHP  & & & & 0.27&0.23&0.27&0.28&0.24&-&-&0.23&0.10&0.38&-&-&0.14&-&0.67&0.31&\textbf{0.282} \\
+		C++  & & & & 0.24&0.12&0.26&0.06&-&0.51&0.35&0.31&0.20&0.26&0.27&0.35&0.73&0.08&0.12&0.43 & \textbf{0.286} \\ 
+		C  & & & & 0.29&0.24&0.29&0.33&0.43&0.56&0.39&0.05&0.03&0.34&0.27&0.33&0.73&0.07&0.23& - & \textbf{0.305}\\
+		C\#  & & & & -&0.20&-&-&0.30&0.89&0.14&0.27&0.60& 0.08 & - & 0.07 & - & 0.57 & - & 0.27 & \textbf{0.339} \\
+		\hline
+		Total Average & &  & & \textbf{0.25}&\textbf{0.17}&\textbf{0.28}&\textbf{0.28}&\textbf{0.28}&\textbf{0.47}&\textbf{0.33}&\textbf{0.17}&\textbf{0.12}&\textbf{0.32}&\textbf{0.27}&\textbf{0.34}&\textbf{0.53}&\textbf{0.11}&\textbf{0.27}&\textbf{0.32} \\
+		\hline
 	\end{tabular}
-  \caption{Blabla}
+	\caption{Ration of hotfix per version update of 7 common languages and their 16 respective domains.
+$\mathcal{P}$ stands for Programming Paradigm class (F=Functional, P=Procedural).
+$\mathcal{C}$ is for Compilation class (S=Static, D=Dynamic), and $\mathcal{T}$ is for Type class (W=weak, S=Strong).}
 	\label{tab:1}
@@
 
-Our work was inspired by the previous research paper \cite{Ray2014}, where authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.
-Our contribution primarily involved investigation of the fault-proneness of different languages based on the newly defined criteria - hotfixes per version update.
-Our dataset consist from [totalnumberofprojects] projects, selected from the top 19 most popular project types in the SourceForge dataset.
-That allows us to differentiate results for the different kinds of projects.\par
-Specifically, our investigation aimed to evaluate the following research questions:
+\section{Results}
+\subsection{RQ1: Language Popularity}
+Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages.
+We identified only 7 widely used languages: C\#, Perl, Python, C, C++ and PHP.
+The criteria of choosing the languages is based on whether the language presents in most of the categories.
+We found that there is a significant drop in the amount of occurrence after Javascript.
+%One would expect Javascript to appear in the list.
+However, it is present only in 7 categories and hence is ignored from the list.
+This may be due to the fact that Javascript is mainly used only for website development while our 16 popular categories span over various aspects of software.
 
-\textbf{1) What languages are the most widely used across the top 16 most popular categories?} (as identified by SourceForge13 Dataset)
-
-Research on the most widely used languages across the top 16 most popular categories showed that there are surprisingly few such languages.
-We identified only 6 widely used languages: C\#, Perl, Python, C, C++ and PHP.
-The criteria of choosing the languages is based on whether the language was present in at least one half of the categories.
-One would expect Javascript to appear in the list.
-However, it is present only in 7 categories which is less than 50\% of the categories.
-This may be due to the fact that Javascript is mainly used for website development.
-The two most widely used languages from the list are C and C++ which are present in 15 out of 16 categories.
-While C++ is not present in site management, C is not present in frameworks.
-The rest of the languages in the list are present on average in 10 categories.
-Overall, the majority of these languages were not present in testing, code generators and site management categories.
+From Table \ref{tab:1}, the two most widely used languages from the list are C and C++ which are present in 15 out of 16 categories.
+While C++ is not present in Site Management, C is not present in Frameworks.
+The rest of the languages in the list are present in 10 categories on average.
+However, it is interesting to note that the majority of these languages are not used in Testing, which contains only Java, C, and C++.
+Similarity, only a few languages appears in Code Generators and Site Management categories.
 The observations made are reasonable since using these languages in such categories is very uncommon.
 
-\textbf{2) What categories achieve the highest/lowest number of hotfixes per version update?}
+\subsection{RQ2: Hotfixes vs Project Categories}
+Next, we investigate which categories achieve the highest or lowest numbers of hotfixes per version update.
+We proceed to analyse why certain categories might have low numbers of hotfixes and vice versa.
 
-In this section we investigate which categories achieve the highest or lowest numbers of hotfixes per version update.
-We proceed to analyse why certain categories might have low numbers of hotfixes and why others might require higher numbers of hotfixes.
+In Table \ref{tab:1}, if we look at the average number of hotfixes per category we see that the Internet, Security, Communications, and Systems Administration have the lowest average number of hot fixes per update out of the entire group of categories.
+The categories with the highest average numbers of hotfixes per version update are Code Generators and Role playing which have 0.53 and 0.47 average hot fixes per update respectively.
+Both these values are much higher than the next highest (0.34).
 
-If we look at the average number of hotfixes per category we see that the Internet, Security, communications and systems administration have the lowest average number of hot fixes per update out of the entire group of categories.
+With regards to those with lower average hot fixes per update, there is a reasonable explanation.
+Applications under the categories as Security, Internet, Communications and Systems administration are all information critical, which means they have critical assets that need to be protected in order to maintain customer's privacy and trust.
+For example, with Communications applications it is vital that security is implemented in such a way that an adversary cannot intercept or obtain messages only intended for the receiver.
+This means that applications that fall under such categories should be extensively tested before deployment to reduce the chance of failure.
+This results in less hot fix being needed after deployment.
 
-The categories with the highest average numbers of hotfixes per version update are Code Generation and Role playing which have 0.53 and 0.47 average hot fixes per update respectively.
-Both these values are much higher than the next highest (0.34) So we will only consider these two.
-
-With regards to those with lower average hot fixes per update it is easy to see why.
-Applications under the categories as Security, Internet, Communications and systems administration are all information critical i.e.
-they have critical assets that need to be protected in order to maintain customer trust and keep the reputation of the creator intact.
-For example with communications applications it is vital that security is implemented in such a way that an adversary cannot intercept or obtain messages only intended for the receiver.
-This means that applications that fall under such categories should be extensively tested before deployment reducing the chance of a hot fix being needed after deployment.
-
-If we look further into the results we can see that for example the usage of C in Internet and security categories has an incredibly low hot fix average (0.03 and 0.05 respectively).
+Another supporting example is the usage of C in Internet and Security categories has an incredibly low hot fix average (0.03 and 0.05 respectively).
 This could be down to fully comprehensive testing techniques, but could also be attributed to the fact that within those categories, the language has been used for an extensive period and hence there is a lot of experience with a variety of different bugs, meaning they are less likely to occur during development as experienced programmers are less likely to make those particular mistakes in the first place.
@@
 
-When considering why role playing applications have a high average number of hot fixes we need to try and speculate as to why the hot fix rate was so much higher than that of projects that fall under the games category.
-It is possible that the games category is a general category are a lot simpler than those that fall under the role playing category which is why bugs might be more prominent as a result of increased complexity of the game.
-If we look a little closer at the results for the role playing category it becomes clear that for all languages there is a high average number of hot fixes per update which supports our theory.
+Role Playing applications have a high average number of hot fixes compared to other projects that fall under the Games category.
+It is possible that the Games category is a general category.
+It is a lot simpler than those that fall under the role playing category which is why bugs might be more prominent as a result of increased complexity of the game.
+Moreover, usually this kind of game contains huge amount of configurations of characters, and items in the game.
+It is common to release a few hotfixes after releasing a new version due to fixing these configurations.
+We can obviously see that all languages in Role Playing category have high average numbers of hot fixes per update.
 
-If we look at Code generators on the other hand we see that for two languages (C and C++) the hot fix average is very high and for two other languages (PHP and Java) it is very low.
-It is not easy to conclude why this might be the case, however one can speculate that PHP and Java are established languages for writing code generators and due to this engender low hot fix averages.
+On the other hand, for Code generators, C and C++, have very high hot fix averages while  PHP and Java are the lowest.
+One can speculate that PHP and Java are established languages for writing code generators and due to this engender low hot fix averages.
 C and C++ might share syntactic details and semantic structure that would result in a similar hot fix average between them.
 
-\textbf{3) What languages achieve the highest/lowest number of hotfixes per version update?}
+\subsection{RQ3: Hotfixes vs Languages}
+Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that Java has appeared to have the least number of hotfixes per version update (0.176) followed by Javascript (0.209) and Perl (0.218).
+However, all the mentioned languages show marginally better results with 8\% difference comparing to other common languages taken under analysis: PHP(0.282), C++(0.286), C (0.305) and C\#(0.339).
+C\# has the highest average ratio of hotfixes among all the languages.
 
-Language comparison based on the average hotfix ratio over all topics has shown that Java has appeared to have the least number of hotfixes per version update (0.18).
-Javascript (0.21) and Perl (0.22) were found to perform slightly worse with the hotfix ratio value being just around 3-4\% higher.
-However, all the mentioned languages showed marginally better results with 8\% difference comparing to other common languages taken under analysis - PHP(0.28), C++(0.29), Python(0.30), C(0.31) and C\#(0.34).
-
-\textbf{4) Are there any patterns based on the type of the programming language?}
-
-The fourth research question that we decided to investigate was related to comparing the type of each programming language.
-The classification was inspired by Ray et al.
-[14], where languages were compared based on the number of bug fixes in each language and were divided into three main Classes.
-The first class was named ‘Programming Paradigm’ and it split languages into the following categories: Procedural, Functional, Scripting.
-The second criteria was Compilation Class, which yielded these types: Static and Dynamic.
-The final one, Type Class, separated languages into: Strongly-typed and weakly-typed.
+\subsection{RQ4: Hotfixes vs Language Types}
+The interesting research question is comparing the type of each programming language by their average ratio of hotfixes.
+The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared based on the number of bug fixes in each language and were divided into three main Classes.
+The first class (Programming Paradigm) has the following categories: Procedural, Functional, Scripting.
+The second class is Compilation containing Static and Dynamic type.
+The final one, Type Class, separates languages into Strongly-typed and weakly-typed.
 
@@
 
-\subsection{Limitations.}
-There were several limitations that we encountered throughout our research:
-\begin{itemize}
-  \item we were able to analyse only a small part of all available projects because most of the developers did not support a versioning system, i.e. no bump commits could be identified.
-  \item we could not extract hotfix branches, only the overall commits on the trunk were available and thus we used our own parameters to try to distinguish hotfix commits.
-Therefore, we may have introduced some false positives or failed to account for actual hotfixes.
-  \item the way we detected bump commits was by writing a regular expression to match words like:  ``bump'', ``version'', ``update'' in an appropriate order.
-Thus, we may have excluded version updates had they been indicated differently.
-  \item most open source projects are not under the stress of releasing as opposed to commercial projects. Consequently, they do not release very often and/or do not support versioning.
-  \item a limited number of programming topics was used (16 in total), which did not allow us to observe distinguishable patterns.
-  \item a low amount of languages satisfied our predicate (being present in at least one half of all evaluated topics), which did not allow us to state any conclusive results.
-  \item only one dataset was used to evaluate our hypothesis - SourceForge 2013.
-\end{itemize}
+\subsection{Limitations}
+There were several limitations that we encountered throughout our study.
+First, we were able to analyse only a small part of all available projects because most of the developers did not support a versioning system, i.e.
+no bump commits could be identified.
+
+%Therefore, we may have introduced some false positives or failed to account for actual hotfixes.
+Moreover, we could not extract hotfix branches, only the overall commits on the trunk were available and thus we used our own parameters to try to distinguish hotfix commits.
+The way we detected bump commits was by writing a regular expression to match words like:  ``bump'', ``version'', ``update'' in an appropriate order.
+Thus, we may generate false positives by including version updates that are not real hotfixes, and also false negatives by excluding version updates had they been indicated differently.
+
+Most open source projects are not under the stress of releasing as opposed to commercial projects. Consequently, they do not release very often and/or do not support versioning.
+
+There is a low amount of languages satisfied our predicate (being present in at least one half of all evaluated topics), which did not allow us to state any conclusive results.
+In addition, with a limited number of programming topics was used (16 in total), we are restricted to observe distinguishable patterns.
+
+Lastly, only one dataset was used to evaluate our hypothesis: SourceForge 2013. Hence, this results may not be generalised.
 

### 2016-02-19 msr_paper/main.tex
Revised from Section 3 onwards until the Conclusion

@@
 	\small
-	\begin{tabular}{c c c c r r r r r r r r r r r r r r r r r}
+	\begin{tabular}{l c c c r r r r r r r r r r r r r r r r r}
 		Language & $\mathcal{P}$ & $\mathcal{C}$ & $\mathcal{T}$ & 
@@
 		\rotatebox{90}{Communications} & \rotatebox{90}{Education} & \rotatebox{90}{Frameworks} &
-		\rotatebox{90}{Total Average} \\
+		\rotatebox{90}{Average} \\
 		\hline
 		Java & P & S & S & 0.25 & 0.15 & 0.22 & 0.08 & 0.28 & 0.32 & 0.07 & 0.22 & 0.38 & 0.17 & 0.23 & 0.10 & 0.06 & 0.09 & 0.10 & 0.09 & \textbf{0.176} \\
-		Javascript  & & & & 0.26 & - & 0.27 & - & 0.17 & - & 0.17 & 0.23 & - & - & - & - & - & - & 0.18 & 0.18 & \textbf{0.209} \\
-		Perl  & & & & 0.17 & 0.05 & 0.22 & 0.44 & - & 0.41 & - & 0.06 & 0.13&0.24&-&-&-&0.17&0.19&0.32 & \textbf{0.218} \\
-		PHP  & & & & 0.27&0.23&0.27&0.28&0.24&-&-&0.23&0.10&0.38&-&-&0.14&-&0.67&0.31&\textbf{0.282} \\
-		C++  & & & & 0.24&0.12&0.26&0.06&-&0.51&0.35&0.31&0.20&0.26&0.27&0.35&0.73&0.08&0.12&0.43 & \textbf{0.286} \\ 
-		C  & & & & 0.29&0.24&0.29&0.33&0.43&0.56&0.39&0.05&0.03&0.34&0.27&0.33&0.73&0.07&0.23& - & \textbf{0.305}\\
-		C\#  & & & & -&0.20&-&-&0.30&0.89&0.14&0.27&0.60& 0.08 & - & 0.07 & - & 0.57 & - & 0.27 & \textbf{0.339} \\
+		C\#  & P & S & S & -&0.20&-&-&0.30&0.89&0.14&0.27&0.60& 0.08 & - & 0.07 & - & 0.57 & - & 0.27 & \textbf{0.339} \\
+	    C++  & P & S & W & 0.24&0.12&0.26&0.06&-&0.51&0.35&0.31&0.20&0.26&0.27&0.35&0.73&0.08&0.12&0.43 & \textbf{0.286} \\ 
+		C  & P & S & W & 0.29&0.24&0.29&0.33&0.43&0.56&0.39&0.05&0.03&0.34&0.27&0.33&0.73&0.07&0.23& - & \textbf{0.305}\\
+		PHP  & S & D & W & 0.27&0.23&0.27&0.28&0.24&-&-&0.23&0.10&0.38&-&-&0.14&-&0.67&0.31&\textbf{0.282} \\		
+		Perl  & S & D & S & 0.17 & 0.05 & 0.22 & 0.44 & - & 0.41 & - & 0.06 & 0.13&0.24&-&-&-&0.17&0.19&0.32 & \textbf{0.218} \\
+		Python	& S & D & S & 0.27	& 0.24	& 0.36	& -	& - & 0.40 & 0.42	& 0.13	& -	& 0.41	& -	& 0.33	& -	& 0.14	& 0.22	& 0.36	& \textbf{0.298} \\
+		JavaScript  & S & D & W & 0.26 & - & 0.27 & - & 0.17 & - & 0.17 & 0.23 & - & - & - & - & - & - & 0.18 & 0.18 & \textbf{0.209} \\
+
 		\hline
-		Total Average & &  & & \textbf{0.25}&\textbf{0.17}&\textbf{0.28}&\textbf{0.28}&\textbf{0.28}&\textbf{0.47}&\textbf{0.33}&\textbf{0.17}&\textbf{0.12}&\textbf{0.32}&\textbf{0.27}&\textbf{0.34}&\textbf{0.53}&\textbf{0.11}&\textbf{0.27}&\textbf{0.32} \\
+		Average & &  & & \textbf{0.25}&\textbf{0.17}&\textbf{0.28}&\textbf{0.28}&\textbf{0.28}&\textbf{0.47}&\textbf{0.33}&\textbf{0.17}&\textbf{0.12}&\textbf{0.32}&\textbf{0.27}&\textbf{0.34}&\textbf{0.53}&\textbf{0.11}&\textbf{0.27}&\textbf{0.32} \\
 		\hline
 	\end{tabular}
-	\caption{Ration of hotfix per version update of 7 common languages and their 16 respective domains.
-$\mathcal{P}$ stands for Programming Paradigm class (F=Functional, P=Procedural).
-$\mathcal{C}$ is for Compilation class (S=Static, D=Dynamic), and $\mathcal{T}$ is for Type class (W=weak, S=Strong).}
+	\caption{Ration of hotfix per version update of 8 common languages and their 16 respective domains.
+$\mathcal{P}$, $\mathcal{C}$, and $\mathcal{T}$ represent the classes of the language.
+$\mathcal{P}$ stands for Programming Paradigm class (P=Procedural, S=Scripting).
+$\mathcal{C}$ is for Compilation class (S=Static, D=Dynamic), and $\mathcal{T}$ is for Type class (W=weakly-typed, S=Strongly-typed).}
 	\label{tab:1}
@@
 \section{Results}
+We performed an empirical study on the SourceForge13 dataset aiming to find the answers for our 4 research questions. The findings of each research question is explained below.
 \subsection{RQ1: Language Popularity}
 Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages.
-We identified only 7 widely used languages: C\#, Perl, Python, C, C++ and PHP.
+We identified only 8 widely used languages: Java, C, C++, Perl, Python, PHP, C\#, and JavaScript.
 The criteria of choosing the languages is based on whether the language presents in most of the categories.
-We found that there is a significant drop in the amount of occurrence after Javascript.
-%One would expect Javascript to appear in the list.
+We found that there is a significant drop in the amount of occurrence after JavaScript.
+%One would expect JavaScript to appear in the list.
 However, it is present only in 7 categories and hence is ignored from the list.
-This may be due to the fact that Javascript is mainly used only for website development while our 16 popular categories span over various aspects of software.
+This may be due to the fact that JavaScript is mainly used only for website development while our 16 popular categories span over various aspects of software.
 
-From Table \ref{tab:1}, the two most widely used languages from the list are C and C++ which are present in 15 out of 16 categories.
+From Table \ref{tab:1}, \textbf{the most widely used languages from the list is Java which appears in all the 16 categories.} C, and C++ share the second place with occurrence in 15 out of 16 categories.
 While C++ is not present in Site Management, C is not present in Frameworks.
@@
 The observations made are reasonable since using these languages in such categories is very uncommon.
+\FIXME{Citation here?}
 
@@
 
-In Table \ref{tab:1}, if we look at the average number of hotfixes per category we see that the Internet, Security, Communications, and Systems Administration have the lowest average number of hot fixes per update out of the entire group of categories.
-The categories with the highest average numbers of hotfixes per version update are Code Generators and Role playing which have 0.53 and 0.47 average hot fixes per update respectively.
+In Table \ref{tab:1}, if we look at the average number of hotfixes per category we see that \textbf{the Internet, Security, Communications, and Systems Administration have the lowest average number of hot fixes per update} out of the entire group of categories.
+The categories with \textbf{the highest average numbers of hotfixes per version update are Code Generators and Role playing} which have 0.53 and 0.47 average hot fixes per update respectively.
 Both these values are much higher than the next highest (0.34).
@@
 \subsection{RQ3: Hotfixes vs Languages}
-Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that Java has appeared to have the least number of hotfixes per version update (0.176) followed by Javascript (0.209) and Perl (0.218).
+Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has appeared to have the least average ratio of hotfixes per version update (0.176)} followed by JavaScript (0.209) and Perl (0.218).
 However, all the mentioned languages show marginally better results with 8\% difference comparing to other common languages taken under analysis: PHP(0.282), C++(0.286), C (0.305) and C\#(0.339).
-C\# has the highest average ratio of hotfixes among all the languages.
+\textbf{C\# has the highest average ratio of hotfixes among all the languages.}
 
@@
 The second class is Compilation containing Static and Dynamic type.
-The final one, Type Class, separates languages into Strongly-typed and weakly-typed.
-
-According to the results that were achieved in the study [14], strongly-typed languages experienced fewer bug fixes than weakly-typed.
+The final one, Type Class, separates languages into Strongly-typed and Weakly-typed.
+According to \cite{Ray2014}, strongly-typed languages experienced fewer bug fixes than its counterpart, weakly-typed.
 Static programming languages required fewer corrective commits than dynamic ones.
@@
 
-Due to the fact that we selected to evaluate languages that appeared in at least 50% of the selected categories, we discovered only 6 languages and none of them was functional.
+Due to our criteria of selecting the languages, the chosen 8 languages are all either procedural or scripting, but not functional.
 Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains.
-Consequently, we evaluated the relationship between Procedural and Scripting languages instead of Procedural against Functional.
+\FIXME{citation here?} Consequently, we compare only Procedural and Scripting languages.
 
-Analysing the first language relationship - Procedural (Java, C, C\#, C++) vs Scripting (Perl, Python, PHP), we were not able to detect any patterns across the categories.
-The reason was that the number of languages was too low to make any firm conclusions.
-It was always the case that at least one of the languages would not follow a pattern or the comparison would consider only one language from a certain category.
+\subsubsection*{Programming Paradigm: Procedural vs Scripting}
+From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among Procedural (Java, C, C\#, C++) and Scripting (Perl, Python, PHP) across all the categories.
+This is possibly because the number of data points (languages) was too low to observe such behaviours and make any firm conclusions.
+%It was always the case that at least one of the languages would not follow a pattern.
+For some categories, the comparison would consider only one language from a certain category.
 
-Evaluating the second relationship between Static (Java, C, C\#, C++) and Dynamic (Perl, Python, PHP) languages led to the same conclusions.
+\subsubsection*{Compilation: Static vs Dynamic}
+In this class, Static (Java, C, C\#, C++) and Dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}.
 The reason was because the languages in these two categories coincided with the languages in the first two classes (Procedural and Scripting).
 
-The third aspect that we investigated referred to comparing strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.
-Similarly to the first language relationship, we encountered that there was no clear pattern across the categories.
-Furthemore, we found the same problem of having comparisons against against only one member of a language class.
-Solely in the Communications category, we were able to detect some pattern, where weakly typed languages performed better.
+\subsubsection*{Type: Strongly-type vs Weakly-typed}
+The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.
+Similarly to the previous two classes, we encountered no clear pattern across the categories.
+Furthermore, we found the same problem of having comparisons against only one member of a language class.
+Solely in the Communications category, we were able to detect some pattern, where strongly-typed languages performed slightly better.
 However, there were only two languages labelled as weak.
@@
 
-Overall, we could not identify any convincing patterns based on the class comparison between languages mostly due to the low number of extracted programming languages.
+Overall, \textbf{we could not identify any convincing patterns based on the class comparison between languages} mostly due to the low number of selected programming languages.
 
@@
 \section{Conclusions and Future Work}
-We have presented a study in detecting the fault-proneness of programming languages through hotfix commits - a topic that has not been investigated so far.
-As part of our research we performed a topic-specific analysis across 16 programming domains to detect the most popular languages used in the evaluated  projects.
-We investigated languages and topics that achieved highest/lowest ratios of hotfix per version update and tried to identify patterns based on the language classification.
-The tool that we created was written in Boa and mined repositories from SourceForge 2013's dataset.
-Our goal was to verify whether the general relationships between languages' reliability holds with regard to hotfixes.
-Our results identified 6 languages as being the most widely used across different domains - C\#, Perl, Python, C, C++ and PHP.
-We discovered that Java achieves the least number of hotfixes per version update (0.18), while C\# requiring most amount of hotfixes per version increase (0.34).
-From the different domains, we recognized: Security, Internet and Communications as being the most reliable (i.e.
-requiring least amount of hotfixes per version update) while Role Playing and Code Generation required highest number of corrections.
-Regarding the comparison between language classes, we were not able to identify any conclusive patterns.\par
-There are few aspects that could be improved as Future Work.
-In particular, more datasets could be used, including Github (provided that projects are categorised into topics).
-Further improvements may be done to identify bump commits more accurately.
-Recently, a few commercial projects have been open-sourced and as a result, we could use them to enhance the quality of our data.
-In such projects, releases will have occurred more frequently and thus, there is a higher probability of identifying more hotfixes.\par
-Finally, we believe that our paper provides a good foundation for future research to be done in this topic.
-We expect that the results we obtained will help developers and project managers improve their planning and understand what the most appropriate languages are for their specific topic.
+We have presented a study in detecting the fault-proneness of programming languages through hotfix commits, a topic that has not been investigated so far.
+As part of our research we performed a topic-specific analysis across 16 programming domains (categories) to detect the most popular languages used in the evaluated  projects.
+We investigated languages and topics that achieved highest/lowest ratios of hotfix per version update and tried to identify patterns based on the language classification presented in \cite{Ray2014}.
+We used Boa to mine repositories in the SourceForge 2013's dataset.
+%Our goal was to verify whether the general relationships between languages' reliability holds with regard to hotfixes.
+We identified 8 languages as being the most widely used across different domains including Java, C, C++, Perl, Python, PHP, C\#, and JavaScript respectively.
+We discovered that Java achieves the least number of hotfixes per version update (0.176), while C\# requiring the highest amount of hotfixes per version increase (0.339).
+From the different categories, we observed that Security, Internet and Communications is the most reliable (i.e.
+requiring least amount of hotfixes per version update) while Role Playing and Code Generation required the highest number of corrections by hotfixes.
+Finally, according to the language classes, we were not able to identify any conclusive patterns.
+
+We believe that there are still aspects to be improved in our study.
+In particular, more datasets could be used, including Github (provided that projects are categorised into topics), and more reliable and accurate method to identify bump commits.
+Instead of using data from SourceForge, real commercial projects which have been open-sourced can be investigated to enhance the quality of our data.
+In such projects, releases should occur more frequently and thus, there is a higher probability of identifying more hotfixes.
+
+Finally, we hope that our paper lays a good foundation for future research to be done in this area.
+We expect that the results we obtained will help developers and project managers improve their planning and understand what are the most appropriate languages for their specific topic or purpose.
 
@@
 
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-%\ifCLASSOPTIONcaptionsoff
-%  \newpage
-%\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-
-%\bibitem{}
-%Hattori, L.P.; Lanza, M., \emph{"On the nature of commits"} in Automated Software Engineering - Workshops, 2008.
-ASE Workshops 2008.
-23rd IEEE/ACM International Conference on , vol., no., pp.63-71, 15-16 Sept.
-2008
-%\bibitem{}
-%Livshits,B.; Zimmerman,T.; \emph{"DynaMine: Finding Common Error Patterns by Mining Software Revision Histories"} in Proceedings of the 10th European software engineering conference held jointly with 13th ACM SIGSOFT international symposium on Foundations of software engineering, pp.296-305
-%\bibitem{}
-%Mockus, A.; Votta, L.G., \emph{"Identifying reasons for software changes using historic databases"} in Software Maintenance, 2000.
-Proceedings.
-International Conference on , vol., no., pp.120-130, 2000
-%\bibitem{}
-%Ratzinger,J.; Sigmund,T.; Gall,H.; \emph{"On the Relation of Refactoring and Software Defects"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.
-35-38
-%\bibitem{}
-%Hindle,A.; German,D.; Holt,R.; \emph{"What do large commits tell us?: a taxonomical study of large commits"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.99-108
-%\bibitem{}
-%Oliva,G.; Steinmacher,I.; Wiese,I.; Aurelio,M.; \emph{"What Can Commit Metadata Tell Us About Design Degradation?"} in IWPSE 2013, Proceedings of the 2013 International Workshop on Principles of Software Evolution, pp.18-27
-%\bibitem{}
-%Phillips,S.; Sillito,J.; Walker,R.; \emph{"Branching and merging: an investigation into current version control practices"} in CHASE '11 Proceedings of the 4th International Workshop on Cooperative and Human Aspects of Software Engineering, pp.9-15
-%\bibitem{}
-%Agrawal, K.; Amreen, S.; Mockus, A., \emph{"Commit Quality in Five High Performance Computing Projects"} in Software Engineering for High Performance Computing in Science (SE4HPCS), 2015 IEEE/ACM 1st International Workshop on , vol., no., pp.24-29, 18-18 May 2015
-%\bibitem{}
-%Steff, M.; Russo, B., \emph{"Co-evolution of logical couplings and commits for defect estimation"} in Mining Software Repositories (MSR), 2012 9th IEEE Working Conference on , vol., no., pp.213-216, 2-3 June 2012
-%\bibitem{}
-%Williams,C.; Spacco,J.; \emph{"Branching and Merging in the Repository"} in MSR '08, Proceedings of the 2008 international working conference on Mining software repositories, pp.19-22
-%\bibitem{}
-%Kolassa,C.; Riehle,D.; Salim,M.; \emph{"The Empirical Commit Frequency Distribution of Open Source Projects"} in WikiSym '13, Proceedings of the 9th International Symposium on Open Collaboration, Article No.18
-%\bibitem{}
-%Ray,B.; Posnett,D.; Filkov,V.; Devanbu,P;  \emph{A Large Scale Study of Programming Languages and Code Quality"} in Github;  FSE 2014, Proceedings of the 22nd ACM SIGSOFT International Symposium on Foundations of Software Engineering, pp.155-165
-%\bibitem{}
-%Phipps,G; \emph{"Comparing Observed Bug and Productivity Rates for Java and C++" } in Software—Practice \& Experience; vol.29, issue 4, pp.345-358, April 10, 1999 
-%\bibitem{}
-%Prechelt, L., \emph{"An empirical comparison of seven programming languages"} in Computer , vol.33, no.10, pp.23-29, Oct 2000
-%\bibitem{}
-%Fourment,M.; Gillings,M.; \emph{"A comparison of common programming languages used in bioinformatics"} in BMC Bioinformatics Feb 2008
-%\bibitem{}
-%Furia,C.; Nanz,S.; \emph{"A Comparative Study of Programming Languages in Rosetta Code"} in ICSE '15, Proceedings of the 37th International Conference on Software Engineering, vol.1, pp.778-788
-%\bibitem{}
-%Iowa State University of Science and Technology,  \emph{"Boa - Mining Ultra-Large-Scale Software Repositories"}, http://boa.cs.iastate.edu/
-
-%\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
 % that's all folks

### 2016-02-19 msr_paper/main.tex
Rewrote abstract, introduction, related work.

--- 2016-02-19 msr_paper/main.tex
@@
-% \documentclass[conference]{IEEEtran}
-
-% \begin{document}
-% \title{An Important Conference Contribution}
-
-% \author{\IEEEauthorblockN{Author One\IEEEauthorrefmark{1},
-% Author Two\IEEEauthorrefmark{2}, Author Three\IEEEauthorrefmark{3} and
-% Author Four\IEEEauthorrefmark{4}}
-% \IEEEauthorblockA{Department of Whatever,
-% Whichever University\\
-% Wherever\\
-% Email: \IEEEauthorrefmark{1}author.one@add.on.net,
-% \IEEEauthorrefmark{2}author.two@add.on.net,
-% \IEEEauthorrefmark{3}author.three@add.on.net,
-% \IEEEauthorrefmark{4}author.four@add.on.net}}
-% \maketitle
-
-% \end{document}
-
 \documentclass{sig-alternate}
@@
 \newcommand\FIXME[1]{\textbf{FIXME: #1}}
-%\fancyhf{}
-%\cfoot{\thepage}
-%\pagestyle{fancy}
 
-% *** GRAPHICS RELATED PACKAGES ***
-%
-%\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-%\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-%\fi
-
-% correct bad hyphenation here
-%\hyphenation{op-tical net-works semi-conduc-tor}
-
-%\pagenumbering{arabic}
 \begin{document}
 \conferenceinfo{MSR}{'16 Austin, Texas USA}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-
-\title{Investigating Fault-Proneness Through Hotfix Commits:\\A Research in the Light of Different Programming Languages}
-\numberofauthors{4}
+\title{Investigating Fault-Proneness Through Hotfix Commits in the Light of Different Programming Languages}
+%\numberofauthors{6}
 \author{
-  Martin Lazarov, Ismaa'eel Ataullah, Vladyslav Kolesnyk, Veselin Pavlov\\
+  Martin Lazarov, Ismaa'eel Ataullah, Vladyslav Kolesnyk, Veselin Pavlov,\\ Chaiyong Ragkhitwetsagul, Jens Krinke\\
   \affaddr{University College London}\\
@@
 
-%\title{Investigating Fault-Proneness Through Hotfix Commits: A Research in the Light of Different Programming Languages}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-%\author{\IEEEauthorblockN{Martin Lazarov\IEEEauthorrefmark{1},
-%Ismaa'eel Ataullah\IEEEauthorrefmark{2}, Vladyslav Kolesnyk\IEEEauthorrefmark{3},
-%Veselin Pavlov\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{Department of Computer Science,
-%UCL\\
-%Email: \IEEEauthorrefmark{1}martin.lazarov.12@ucl.ac.uk,
-%\IEEEauthorrefmark{2}ismaaeel.ataullah.12@ucl.ac.uk,
-%\IEEEauthorrefmark{3}vladyslav.kolesnyk.12@ucl.ac.uk,
-%\IEEEauthorrefmark{4}veselin.pavlov.15@ucl.ac.uk}}
-
-% make the title area
 \maketitle
@@
 \begin{abstract}
-%\boldmath
-Choosing the most suitable programming language is a very challenging task for every project manager and developer.
+Choosing the most suitable programming language is a challenging task for every project manager and developer.
 Often the success of a software project is determined by the stability of the system after a release.
-In this paper, we propose analysing hotfix commits in various development areas as a means of indicating what languages tend to produce the least amount of problems after software has been published.
-We have focused on parsing metadata from the 16 most popular projects hosted on SourceForge and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages.
+We analyse the prevalence of hotfix commits in various development areas as an indicator of which languages tend to produce the least amount of problems after software has been published.
+Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are use across project types.
+Our results only identified 6 languages that are widely used across different domains, Java, C, C++, C\#, Perl, and PHP.
+Java has the least number of hotfixes per version update (0.18), while C\# required most hotfixes per version update (0.34).
+We identified the Security and Internet \& Communications project types as being the most reliable (i.e.\ requiring least amount of hotfixes per version update).
+The analysis did not reveal conclusive patterns for language classes (strongly vs.\ weakly typed and dynamic vs.\ static).
 \end{abstract}
@@
 % A category with the (minimum) three required fields
-\category{D.2.7}{Software Engineering}{\FIXME{}}
+%\category{D.2.7}{Software Engineering}{\FIXME{}}
 %A category including the fourth, optional field follows...
@@
 
-\keywords{Version Control, hotfixes, SourceForge, commits metadata}
-
-%\begin{IEEEkeywords}
-%Version Control, hotfixes, SourceForge, commits metadata.
-%\end{IEEEkeywords}
+%\keywords{Version Control, hotfixes, SourceForge, commits metadata}
 
@@
 Planning and resource allocation are arguably the most crucial components of any software project.
-Our research was motivated by the fact that very often, in the planning phase project managers neglect the importance of hotfixes - which becomes essential once a project is deployed.
-
-Generally, a hotfix is related to correcting a critical problem which prevents a particular software product from performing as expected.
+In the planning phase, project managers often neglect the importance of hotfixes -- which becomes essential once a project is deployed.
+A hotfix is correcting a critical problem which prevents a particular software product from performing as expected.
 This could mean a total or partial loss of service, a major security breach that has to be patched instantly, etc.
 As a result, such a fix is done outside of normal development and testing procedures.
-This characteristic of hotfixes implies that rapid development under urgent circumstances is usually required.
-Therefore, it can be often the case that lots of additional people and resources need to be allocated quickly for the correction, which leads to a significant increase in the cost of the overall product.
+The characteristic of hotfixes implies that rapid development under urgent circumstances is usually required.
+Therefore, it can be often the case that additional resources need to be allocated quickly for the correction, which leads to a significant increase in the cost of the overall product.
 
-Another aspect that makes hotfixes so important is their unpredictability which further prevents teams from planning in advance for such corrections into their development cycle.
+% Another aspect that makes hotfixes so important is their unpredictability which further prevents teams from planning in advance for such corrections into their development cycle.
 Moreover, hotfixes are sometimes delayed on purpose because people are afraid of the risk of applying the fix.
@@
 
-As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: ``a corrective commit deployed within 1 day, following a project's version update to fix critical issues in the release''.\par
+There are no clear definitions of what a hotfix is, we define a hotfix as \emph{a corrective commit deployed shortly after a project's release to fix a critical issue}.
+Inspired by work on the impact of programming languages on code quality, we investigated how programming languages impact the frequency of hotfixes.
+Because project domains strongly influence the choice of a programming language, we restricted our analysis on programming languages that are used \emph{across} project domains.
 
-With our research we aim to help project managers and developers make a more informed decision when deciding what languages to use by comparing the number of hotfixes per programming language.
+%With our research we aim to help project managers and developers make a more informed decision when deciding what languages to use by comparing the number of hotfixes per programming language.
 To provide a better picture we also look at hotfixes per specific software topics.
 
-We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used.
+%We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used.
 Consequently, project managers will be able to make a more informed decision with regard to the most appropriate programming languages they could use to try to achieve a minimum number of problems in production.
@@
 Furthermore, it will help optimise the allocation of resources.
-The contributions of this paper are:
 
+The main contributions of this paper are:
+%
 \begin{itemize}
-  \item parameters to determine most probable hotfix commits in the repository.
-  \item a quantitative analysis identifying fault-proneness of specific programming topics and the relevant distribution of faults across programming languages.
-  \item an evaluation of the most used languages used across the top 16 most popular domains identified by SourceForge 13'.
-  \item an investigation into the languages that achieve the highest/lowest ratios of hotfix per version update.
-  \item research into the programming topics that experience the highest/lowest numbers of hotfix per version increase.
-  \item an investigation into the following relationships between languages: strongly-typed vs weakly-typed, static vs dynamic and scripting vs procedural to try to determine what group of languages is the most reliable.
-  \item a foundation for future research in the area of identifying language-specific hotfixes.
+  \item An  approach to identify hotfix commits in repositories.
+  \item A quantitative analysis identifying fault-proneness of specific programming topics and the distribution of faults across programming languages.
+The analysis identified that only six languages are used across the top 16 most popular domains identified by the Boa SourceForge 2013 dataset.
+  \item An investigation on the impact of language classes on reliability: strongly-typed vs weakly-typed, static vs dynamic and scripting vs procedural.
+The investigation showed no clear pattern on the impact of language classes.
 \end{itemize}
-
-The rest of the paper is organised as follows: Related work is discussed in section 2.
-Section 3 explains our empirical study with 4 research questions.
-In Section 4, the results are presented and discussed and followed by our conclusion and future work in Section 5.
+%
+Related work is discussed in the next section. Section 3 explains our empirical study and its results, followed by conclusions and future work.
 
 \section{Related Work}
-Many researchers have explored various aspects of defects in software through the use of versioning system history logs.
+%Many researchers have explored various aspects of defects in software through the use of versioning system history logs.
 We have also decided to follow this approach due to its openness and accessibility.
 This method is also very widely adopted amongst the MSR community \cite{Hattori2008}.
-Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages' stability.
+
+Our investigation focused primarily on two research aspects and existing work related to 1) comparison of programming languages impact, and 2) identifying hotfix commits.
+
+% Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed \cite{Oliva2013}.
+However, they did not make their investigation language-specific as opposed to what we propose in our paper.\par
+
+% Other scientists tried to explore the logical coupling of commits as a way of identifying potential software defects.
+They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future \cite{Steff2012}.\par
+
+% A further interesting research that we found was related to the use of separate bug fix branches.
+According to the study performed by Phillips et al.
+\cite{Phillips2011}, 41\% of the tested projects had incorporated an isolated bug fix branch.
+
+\subsection{Comparison of programming languages}
+%Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality \cite{Agrawal2015}.
+Furthermore, there were several papers that compared and contrasted the performance of different languages, which led to our second research topic.
+We explored these existing scientific studies to see what techniques authors had implemented.
+We also analysed the hypotheses these papers proposed to increase the scope of our project.\par
+
+%Our specific point of interest was related to different software topics and programming languages to discover how number of hotfixes per topic depend on the language developers use.\par
+
+The main related work focussing on the effect of programming languages~\cite{Ray2014} is based on the rate of defect occurrences in various languages.
+The study was carried out on a huge dataset from GitHub (1.5 million commits).
+To evaluate the effect that programming languages have on software quality, the authors grouped languages according to their type, e.g.\ strongly vs.\ weakly typed.
+They also used mathematical methods to triangulate and produce statistics for how group size, project history and number of contributors affect the number of defects.
+In order to achieve more representative results, they took a sample of 17 languages and extracted the top 50 projects per language.
+The results showed how defect-prone languages are compared to others.
+This research focused on classifying bugs based on their type, while we focused specifically on the distribution of hotfixes across programming languages.
+
+Phipps et al.~\cite{Phipps1999} evaluated Java and C++ by comparing bug rates.
+There were two definitions for bugs: defect and bug.
+Defects included issues such as syntax errors, logical errors and changes in variable name.
+A bug itself was only defined to be a problem detected during testing or deployment.
+The mean number of defects and bugs, averaged over a number of projects, was compared between both languages per 1000 lines of code.
+The work also observed defects and bugs per hour as well as the time taken to fix defects in both languages.
+Productivity was measured by analysing number of lines written per second.
+
+Other work used different parameters to compare programming languages, for example, program length, programming effort, run-time efficiency, memory consumption, and reliability \cite{Prechelt2000}.
+Fourment and Gilling~\cite{Fourment2008} analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages.
+This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.
+
+Another recent work~\cite{Nanz2015} compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness, arguing that conciseness is an important factor as a more compact language can help write a program with fewer bugs.
 
 \subsection{Identifying hotfix commits}
-The first major problem we had to solve was identifying actual hotfix commits.
+%The first major problem we had to solve was identifying actual hotfix commits.
 Due to Boa's limitations, we were not able to find commit history from hotfix branch.
@@
 
-A large number of the reports that we studied implemented a very simple approach.
-They focused on the commit message as the main mechanism to try to identify whether a commit was of a corrective type~\cite{Hattori2008, Ratzinger2008, Hindle2008}.
-They simply investigated if the string of text contained particular key words, such as: ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.\par
+A large number of previous work use a simple approach to identify bug-fixing commits.
+They use the commit message as the main mechanism to identify whether a commit was of a corrective type by simply investigating if the message contains particular key words, such as ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.~\cite{Hattori2008, Ratzinger2008, Hindle2008}.
+Others proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages.
+Their algorithm consists of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g.
+``for'', ``code'', ``to'', ``the'', etc.
+) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits~\cite{Mockus2000}.
 
-Some authors proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages.
-Their algorithm consisted of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g.
-``for'', ``code'', ``to'', ``the'', etc.
-) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits \cite{Mockus2000}.\par
+Another approach to identify bug fix commits performs a classification based on the size of the commits~\cite{Hattori2008,Mockus2000,Hindle2008}.
+The size of a commit can be measured based on the number of lines or the amount of files that were changed.
+Large commits tend to improve code quality or add new features, while small ones are considered to be corrective -- targeted at fixing bugs in the project.
+A fine-grained size classification depending on the number of files changed per commit defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large~\cite{Hattori2008}.
+The authors argued that most of the commits tend to be tiny (in 80\% of the cases) which implies that bug fix commits occur very often as corrective commits tend to be small ones.
 
-Another approach to identify bug fix commits involved performing a classification based on the size of the commits \cite{Hattori2008,Mockus2000,Hindle2008}.
-The size of a commit can be measured based on the number of lines or the amount of files that were changed.
-According to these authors, large commits tend to improve code quality or add new features, while small ones are considered to be corrective - targeted at fixing bugs in the project.
-Some researchers proposed a further fine-grained size classification depending on the number of files changed per commit \cite{Hattori2008}.
-They defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large (126 up ).
-The scientists further argued that most of the commits tend to be tiny (in 80\% of the cases).
-This implies that bug fix commits occur very often as corrective commits tend to be small ones.\par
+Issues in critical components (e.g.~errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered~\cite{Livshits2005}.
+Mockus and Votta~\cite{Mockus2000} argue that the time between changes will depend on the type of alteration and that corrective changes require the least amount of time to be completed as they often have to be performed promptly regardless of the complexity of the task.
 
-We also looked at the time factor with regard to commits.
-According to some authors, issues in critical components (e.g.
-errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered \cite{Livshits2005}.
-In \cite{Mockus2000}, the researchers argue that the time between changes will depend on the type of alteration.
-Furthermore, they hold the view that corrective changes require the least amount of time to be completed as often they have to be performed promptly regardless of the complexity of the task.\par
-
-In terms of the time aspect, there was another paper which investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk \cite{Williams2008}.
+Williams and Spacco~\cite{Williams2008} investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk.
 However, they explored branches in general without focusing specifically on bugfix branches.
-Some other scientists were trying to explain people's behaviour during commits by investigating their commit frequency distribution.
-As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits \cite{Kolassa2013}.\par
-
-Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed \cite{Oliva2013}.
-However, they did not make their investigation language-specific as opposed to what we propose in our paper.\par
-
-Other scientists tried to explore the logical coupling of commits as a way of identifying potential software defects.
-They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future \cite{Steff2012}.\par
-
-A further interesting research that we found was related to the use of separate bug fix branches.
-According to the study performed by Phillips et al.
-\cite{Phillips2011}, 41\% of the tested projects had incorporated an isolated bug fix branch.
-
-\subsection{Comparison of programming languages' stability}
-Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality \cite{Agrawal2015}.
-Furthermore, there were several papers that compared and contrasted the performance of different languages, which led to our second research topic.
-We explored these existing scientific studies to see what techniques authors had implemented.
-We also analysed the hypotheses these papers proposed to increase the scope of our project.\par
-
-Our specific point of interest was related to different software topics and programming languages to discover how number of hotfixes per topic depend on the language developers use.\par
-
-There were researchers that focused on the effect of programming languages on software quality \cite{Ray2014}.
-The authors based performance on the rate of defect occurrences in various languages.
-The study was carried out on a huge dataset from GitHub (1.5 million commits).
-To evaluate the effect that programming languages have on software quality, the authors implemented the following strategy.
-First, they grouped languages according to their type, i.e.
-strongly vs weakly typed.
-They also used mathematical methods to triangulate and produce statistics for how group size, project history and number of contributors affect the number of defects.
-In order to achieve more representative results, they took a sample of 17 languages and extracted the top 50 projects per language.
-The results showed how defect-prone languages are compared to others.
-This was done by taking a grand mean and then comparing each language to the grand mean.
-Languages were then grouped, a mean was taken and then types of languages were once again compared to a grand mean - e.g.
-functional-strong-static languages compared to functional-dynamic-strong languages.
-This research focused on classifying bugs based on their type as opposed to our investigation, where we focused specifically on the distribution of hotfixes across programming languages.\par
-
-In \cite{Phipps1999}, Java and C++ were evaluated by comparing bug rates.
-In this case, there were two definitions for bugs - a defect and a bug itself.
-A defect included issues such as syntax errors, logical errors and changes in variable name.
-A bug itself was only defined to be a problem detected during testing or deployment.
-The mean number of defects and bugs, averaged over a number of projects, was compared between both languages per 1000 lines of code.
-The scientists also compared observed defects and bugs per hour as well as the time taken to fix defects in both languages.
-Productivity was also measured by analysing number of lines written per second.\par
-
-There were researchers that used slightly different parameters to compare programming languages, these included program length, programming effort, run-time efficiency, memory consumption, and reliability \cite{Prechelt2000}.\par
-
-An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages \cite{Fourment2008}.
-This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.\par
-
-Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness \cite{Nanz2015}.
-They argued that conciseness is an important factor as a more compact language can help write a program with fewer bugs.
+Kolassa et al.~\cite{Kolassa2013} try to explain people's behaviour during commits by investigating their commit frequency distribution.
+They discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits.
 

### 2016-02-19 msr_paper/main.tex
fixing FIXMEs and adding a paper to references

@@
 The SourceForge13 Dataset is selected for this study.
-It consists of \FIXME{[totalnumberofprojects]} projects, selected from the top 16 most popular project types in the SourceForge dataset.
+It consists of \textbf{49865} projects, selected from the top 16 most popular project types in the SourceForge dataset.
 That enables us to differentiate results for the different kinds of projects.
@@
 Similarity, only a few languages appears in Code Generators and Site Management categories.
-The observations made are reasonable since using these languages in such categories is very uncommon.
-\FIXME{Citation here?}
 
@@
 Due to our criteria of selecting the languages, the chosen 8 languages are all either procedural or scripting, but not functional.
-Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains.
-\FIXME{citation here?} Consequently, we compare only Procedural and Scripting languages.
+Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains  \cite{Wadler1999}.
+Consequently, we compare only Procedural and Scripting languages.
 

### 2016-02-20 msr_paper/main.tex
Fixed the number of programming languages in the abstract from 6 to 8
Fixed some typos and turning all headers into lower case

@@
 Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are use across project types.
-Our results only identified 6 languages that are widely used across different domains, Java, C, C++, C\#, Perl, and PHP.
+Our results only identified 8 languages that are widely used across different domains, Java, C, C++, C\#, Perl, Python, PHP, and JavaScript.
 Java has the least number of hotfixes per version update (0.18), while C\# required most hotfixes per version update (0.34).
@@
 
-\subsection{Identifying hotfix commits}
+\subsection{Identifying bug fix commits}
 %The first major problem we had to solve was identifying actual hotfix commits.
@@
 
-\subsection{Research Questions}
+\subsection{Approach for identifying hotfix commits}
+
+\subsection{Research questions}
 \noindent Our study aimed to answer the following research questions:
 
-\noindent\textbf{RQ1 (Language Popularity):} What languages are the most widely used across the top 16 most popular categories?
+\noindent\textbf{RQ1 (language popularity):} What languages are the most widely used across the top 16 most popular categories?
 
-\noindent\textbf{RQ2 (Hotfixes vs Project Categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ2 (hotfixes vs project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ3 (Hotfixes vs Languages):} What languages achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ3 (hotfixes vs languages):} What languages achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ4 (Hotfixes vs Language Types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?
+\noindent\textbf{RQ4 (hotfixes vs language types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?
 
@@
 We performed an empirical study on the SourceForge13 dataset aiming to find the answers for our 4 research questions. The findings of each research question is explained below.
-\subsection{RQ1: Language Popularity}
+\subsection{RQ1: language popularity}
 Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages.
@@
 
-\subsection{RQ2: Hotfixes vs Project Categories}
+\subsection{RQ2: hotfixes vs project categories}
 Next, we investigate which categories achieve the highest or lowest numbers of hotfixes per version update.
@@
 
-\subsection{RQ3: Hotfixes vs Languages}
+\subsection{RQ3: hotfixes vs languages}
 Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has appeared to have the least average ratio of hotfixes per version update (0.176)} followed by JavaScript (0.209) and Perl (0.218).
@@
 
-\subsection{RQ4: Hotfixes vs Language Types}
-The interesting research question is comparing the type of each programming language by their average ratio of hotfixes.
+\subsection{RQ4: hotfixes vs language types}
+Another interesting research question is comparing the type of each programming language by their average ratio of hotfixes.
 The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared based on the number of bug fixes in each language and were divided into three main Classes.
@@
 
-\subsubsection*{Programming Paradigm: Procedural vs Scripting}
+\subsubsection*{Programming paradigm: procedural vs scripting}
 From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among Procedural (Java, C, C\#, C++) and Scripting (Perl, Python, PHP) across all the categories.
@@
 
-\subsubsection*{Compilation: Static vs Dynamic}
-In this class, Static (Java, C, C\#, C++) and Dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}.
-The reason was because the languages in these two categories coincided with the languages in the first two classes (Procedural and Scripting).
+\subsubsection*{Compilation: static vs dynamic}
+In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}.
+The reason was because the languages in these two categories coincided with the languages in the first two classes (procedural and scripting).
 
-\subsubsection*{Type: Strongly-type vs Weakly-typed}
+\subsubsection*{Type: strongly-type vs weakly-typed}
 The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.

### 2016-02-20 msr_paper/main.tex
Revised on section 3-5 by adding the method to identify hotfixes

@@
 \subsection{Dataset}
-The SourceForge13 Dataset is selected for this study.
-It consists of \textbf{49865} projects, selected from the top 16 most popular project types in the SourceForge dataset.
-That enables us to differentiate results for the different kinds of projects.
+The Boa September 2013 SourceFourge dataset is selected for this study.
+It consists of 49,865 projects, selected from the top 16 most popular project categories in the dataset.
+We only concentrate on projects that incorporate versioning (i.e.
+containing bump commits).
+%That enables us to differentiate results of each different project category.
 
 \subsection{Approach for identifying hotfix commits}
+We invented a method to distinguish a hotfix based on commit messages. The method has 2 parts:
+
+\textbf{1) Detecting version bump commits:} From the observed data, it turned out to be straightforward to extract this information by a regular expression that matches the following strings: ``bump'', ``version'', ``update'' in an appropriate order.
+
+\textbf{2) Detecting hotfix commits:} Due to the similarity between commit messages for normal bug fixes and hotfix commits, it was challenging to distinguish them only based on the message.
+Therefore, we applied two additional filters: time and range of commits.
+
+\textbf{Time} -- hotfix commits have to be submitted within 1 day from the version bump commit.
+Thus, a hotfix is assigned the highest priority and has to be submitted rapidly.
+Furthermore, it is very likely that fixes performed on the first day after a new release will refer to problems occurring before the version increase.
+
+\textbf{Look-ahead range for commits} -- only n numbers of commits should be considered after a version bump.
+We focused on the Look-ahead parameter and decided to investigate only the first 10 commits as any commit after this point would possibly be related to code added after the release rather than before that.
+We have performed an initial investigation of 5, 10 and 15 commits and found that 10 achieved the best balance between high number of fix commits and low number of commits that introduce new features.
 
@@
 
-\noindent\textbf{RQ2 (hotfixes vs project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ2 (hotfixes vs. project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ3 (hotfixes vs languages):} What languages achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ3 (hotfixes vs. languages):} What languages achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ4 (hotfixes vs language types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?
+\noindent\textbf{RQ4 (hotfixes vs. language types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?
 
@@
 	\end{tabular}
-	\caption{Ration of hotfix per version update of 8 common languages and their 16 respective domains.
+	\caption{Ratios of hotfix commits per version bump commits of 8 common languages and their 16 respective categories.
 $\mathcal{P}$, $\mathcal{C}$, and $\mathcal{T}$ represent the classes of the language.
@@
 \section{Results}
-We performed an empirical study on the SourceForge13 dataset aiming to find the answers for our 4 research questions. The findings of each research question is explained below.
+We performed an empirical study on the SourceForge dataset using our method of identifying version bump commits and hotfix commits. The findings are explained below.
 \subsection{RQ1: language popularity}
 Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages.
-We identified only 8 widely used languages: Java, C, C++, Perl, Python, PHP, C\#, and JavaScript.
-The criteria of choosing the languages is based on whether the language presents in most of the categories.
-We found that there is a significant drop in the amount of occurrence after JavaScript.
+We chose the languages based on their frequency of occurrence across 16 categories.
+From Table \ref{tab:1}, the 8 widely used languages include Java (16), C (15), C++ (15), Perl (11), Python (11), PHP (11), C\# (10), and JavaScript (7) respectively.
+We found a significant drop in occurrence for the rest of languages so their results are not included here.
 %One would expect JavaScript to appear in the list.
@@
 
-From Table \ref{tab:1}, \textbf{the most widely used languages from the list is Java which appears in all the 16 categories.} C, and C++ share the second place with occurrence in 15 out of 16 categories.
+It is interesting that majority of these languages are not used in Testing, which contains only Java, C, and C++.
+Similarity, only a few languages appears in Code Generators and Site Management categories.
+
+% From Table \ref{tab:1}, \textbf{the most widely used languages from the list is Java which appears in all the 16 categories.} C, and C++ share the second place with occurrence in 15 out of 16 categories.
 While C++ is not present in Site Management, C is not present in Frameworks.
@@
 
-\subsection{RQ2: hotfixes vs project categories}
-Next, we investigate which categories achieve the highest or lowest numbers of hotfixes per version update.
-We proceed to analyse why certain categories might have low numbers of hotfixes and vice versa.
+\subsection{RQ2: hotfixes vs. project categories}
+To be able to compare among languages, we calculated their hotfix ratios.
+The hotfix ratio of a language $P$ in category $C$ is a ratio of total hotfix commits per total version bump commits of the language in that category.
+%We proceed to analyse why certain categories might have low numbers of hotfixes and vice versa.
+\begin{equation*}
+\mathrm{HotfixRatio}_{P,C} = \frac{\sum{\mathrm{HotfixCommit}_{P,C}}}{\sum{\mathrm{VersionBumpCommit}_{P,C}}}
+\end{equation*}
+In Table \ref{tab:1}, \textbf{Communications have the lowest average hotfix ratios across all categories (0.11)} followed by Internet, Security, and Systems Administration (0.12, 0.17, 0.17) respectively.
+These categories are information critical.
+They are expected to be tested extensively before deployment to reduce the chance of failure.
+This may results in less hot fixes being needed after deployment.
+An interesting observation is that the average hotfix ratio of C in Internet category is incredibly low (0.03).
+We believe that is because of there are several new languages designed specifically for Internet applications.
+Not many new C code are being written in this area which results in a very low hotfix commits.
+%Security categories  %This could be down to fully comprehensive testing techniques, but could also be attributed to the fact that within those categories, the language has been used for an extensive period and hence there is a lot of experience with a variety of different bugs, meaning they are less likely to occur during development as experienced programmers are less likely to make those particular mistakes in the first place.
+This could also be the reason the projects written in Perl have a very low hot fix average for all four categories.
 
-In Table \ref{tab:1}, if we look at the average number of hotfixes per category we see that \textbf{the Internet, Security, Communications, and Systems Administration have the lowest average number of hot fixes per update} out of the entire group of categories.
-The categories with \textbf{the highest average numbers of hotfixes per version update are Code Generators and Role playing} which have 0.53 and 0.47 average hot fixes per update respectively.
-Both these values are much higher than the next highest (0.34).
+On the other hand, the categories with \textbf{the highest average hotfix ratios are Code Generators (0.53)}.
+C and C++, have very high average hot fix ratios while  PHP and Java are the lowest.
+One can speculate that PHP and Java are established languages for writing code generators and due to this engender low hot fix averages.
+C and C++ might share syntactic details and semantic structure that would result in a similar hot fix average between them.
 
-With regards to those with lower average hot fixes per update, there is a reasonable explanation.
-Applications under the categories as Security, Internet, Communications and Systems administration are all information critical, which means they have critical assets that need to be protected in order to maintain customer's privacy and trust.
+The second place is Role playing category (0.47) which has higher hotfix ratio than its superset, Game category.
+We believe that this specific kind of game is more complex with vast amount of configurations of characters and items in the game.
+It is common to release a few hotfixes after a new version release due to adjusting these configurations.
+%We can obviously see that all languages in Role Playing category have high average numbers of hot fixes per update.
+
+%Applications under the categories as Security, Internet, Communications and Systems administration are all information critical, which means they have critical assets that need to be protected in order to maintain customer's privacy and trust.
 For example, with Communications applications it is vital that security is implemented in such a way that an adversary cannot intercept or obtain messages only intended for the receiver.
@@
 
-Another supporting example is the usage of C in Internet and Security categories has an incredibly low hot fix average (0.03 and 0.05 respectively).
-This could be down to fully comprehensive testing techniques, but could also be attributed to the fact that within those categories, the language has been used for an extensive period and hence there is a lot of experience with a variety of different bugs, meaning they are less likely to occur during development as experienced programmers are less likely to make those particular mistakes in the first place.
-This could also be the reason the projects written in Perl have a very low hot fix average for all four categories.
-
-Role Playing applications have a high average number of hot fixes compared to other projects that fall under the Games category.
+%Role Playing applications have a high average number of hot fixes compared to other projects that fall under the Games category.
 It is possible that the Games category is a general category.
@@
 
-On the other hand, for Code generators, C and C++, have very high hot fix averages while  PHP and Java are the lowest.
+%On the other hand, for Code generators, C and C++, have very high hot fix averages while  PHP and Java are the lowest.
 One can speculate that PHP and Java are established languages for writing code generators and due to this engender low hot fix averages.
@@
 
-\subsection{RQ3: hotfixes vs languages}
-Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has appeared to have the least average ratio of hotfixes per version update (0.176)} followed by JavaScript (0.209) and Perl (0.218).
-However, all the mentioned languages show marginally better results with 8\% difference comparing to other common languages taken under analysis: PHP(0.282), C++(0.286), C (0.305) and C\#(0.339).
-\textbf{C\# has the highest average ratio of hotfixes among all the languages.}
+\subsection{RQ3: hotfixes vs. languages}
+Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has the least average hotfix ratio (0.176)} followed by JavaScript (0.209) and Perl (0.218).
+However, all the mentioned languages show marginally better results with approximately 8\% difference comparing to other common languages: PHP (0.282), C++ (0.286), C (0.305) and C\#(0.339).
+\textbf{C\# has the highest average hotfix ratio among all the languages.}
 
-\subsection{RQ4: hotfixes vs language types}
-Another interesting research question is comparing the type of each programming language by their average ratio of hotfixes.
-The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared based on the number of bug fixes in each language and were divided into three main Classes.
-The first class (Programming Paradigm) has the following categories: Procedural, Functional, Scripting.
-The second class is Compilation containing Static and Dynamic type.
-The final one, Type Class, separates languages into Strongly-typed and Weakly-typed.
+\subsection{RQ4: hotfixes vs. language types}
+Another interesting research question is comparing the type of each programming language by their average hotfix ratio.
+The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared using hotfix ratios with respect to their three main classes.
+Programming Paradigm class (noted as $\mathcal{P}$ in Table \ref{tab:1}) has three variations: procedural, functional, and scripting.
+The Compilation class contains static and dynamic type.
+The final one, Type class, separates languages into strongly-typed and weakly-typed.
 According to \cite{Ray2014}, strongly-typed languages experienced fewer bug fixes than its counterpart, weakly-typed.
@@
 Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains  \cite{Wadler1999}.
-Consequently, we compare only Procedural and Scripting languages.
+Consequently, we only compare procedural and scripting languages in this class.
 
-\subsubsection*{Programming paradigm: procedural vs scripting}
-From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among Procedural (Java, C, C\#, C++) and Scripting (Perl, Python, PHP) across all the categories.
-This is possibly because the number of data points (languages) was too low to observe such behaviours and make any firm conclusions.
+\subsubsection*{Programming paradigm: procedural vs. scripting}
+From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among procedural (Java, C, C\#, C++) and scripting (Perl, Python, PHP) across all the categories.
+This is possibly because the number of data points (languages) were too low to observe such behaviours and make any firm conclusions.
 %It was always the case that at least one of the languages would not follow a pattern.
@@
 
-\subsubsection*{Compilation: static vs dynamic}
+\subsubsection*{Compilation: static vs. dynamic}
 In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}.
@@
 
-\subsubsection*{Type: strongly-type vs weakly-typed}
+\subsubsection*{Type: strongly-type vs. weakly-typed}
 The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.
@@
 
-We believe that there are still aspects to be improved in our study.
+We believe that there are still some aspects to be improved in our study.
 In particular, more datasets could be used, including Github (provided that projects are categorised into topics), and more reliable and accurate method to identify bump commits.

### 2016-02-21 msr_paper/main.tex
Martin and Vlad small changes. Including referencing in introduction that that we have chosen 8 languages instead of 6.

@@
 We analyse the prevalence of hotfix commits in various development areas as an indicator of which languages tend to produce the least amount of problems after software has been published.
-Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are use across project types.
-Our results only identified 8 languages that are widely used across different domains, Java, C, C++, C\#, Perl, Python, PHP, and JavaScript.
+Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are used across project types.
+Our results only identified 8 languages that are widely used across different domains: Java, C, C++, C\#, Perl, Python, PHP, and JavaScript.
 Java has the least number of hotfixes per version update (0.18), while C\# required most hotfixes per version update (0.34).
@@
 There are no clear definitions of what a hotfix is, we define a hotfix as \emph{a corrective commit deployed shortly after a project's release to fix a critical issue}.
-Inspired by work on the impact of programming languages on code quality, we investigated how programming languages impact the frequency of hotfixes.
-Because project domains strongly influence the choice of a programming language, we restricted our analysis on programming languages that are used \emph{across} project domains.
+Inspired by work on the impact of programming languages on code quality, we investigated how programming languages affect the frequency of hotfixes.
+Because project domains strongly influence the choice of a programming language, we restricted our analysis to programming languages that are used \emph{across} project domains.
 
@@
   \item A quantitative analysis identifying fault-proneness of specific programming topics and the distribution of faults across programming languages.
-The analysis identified that only six languages are used across the top 16 most popular domains identified by the Boa SourceForge 2013 dataset.
+The analysis identified that only eight languages are used across the top 16 most popular domains identified by the Boa SourceForge 2013 dataset.
   \item An investigation on the impact of language classes on reliability: strongly-typed vs weakly-typed, static vs dynamic and scripting vs procedural.
-The investigation showed no clear pattern on the impact of language classes.
+The investigation showed no clear patterns on the influence of language classes.
 \end{itemize}
@@
 
-A large number of previous work use a simple approach to identify bug-fixing commits.
-They use the commit message as the main mechanism to identify whether a commit was of a corrective type by simply investigating if the message contains particular key words, such as ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.~\cite{Hattori2008, Ratzinger2008, Hindle2008}.
-Others proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages.
-Their algorithm consists of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g.
+A large number of previous work uses a simple approach to identify bug-fixing commits.
+The commit message was employed as the main mechanism to identify whether a commit was of a corrective type by simply investigating if the message contained particular key words, such as ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.~\cite{Hattori2008, Ratzinger2008, Hindle2008}.
+Other authors proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages.
+Their algorithm consisted of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g.
 ``for'', ``code'', ``to'', ``the'', etc.
@@
 Large commits tend to improve code quality or add new features, while small ones are considered to be corrective -- targeted at fixing bugs in the project.
-A fine-grained size classification depending on the number of files changed per commit defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large~\cite{Hattori2008}.
+A fine-grained size classification depending on the number of files changed per commit defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), 4) large (> 125 files) ~\cite{Hattori2008}.
 The authors argued that most of the commits tend to be tiny (in 80\% of the cases) which implies that bug fix commits occur very often as corrective commits tend to be small ones.
@@
 
-Our work was inspired by the previous research paper \cite{Ray2014}, where authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.
+Our work was inspired by a research paper \cite{Ray2014}, where authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.
 Our contribution primarily involved investigation of the fault-proneness of different languages based on the newly defined criteria - hotfixes per version update.
@@
 
-\noindent\textbf{RQ4 (hotfixes vs. language types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?
+\noindent\textbf{RQ4 (hotfixes vs. language types):} Are there any relationships between the occurrences of hotfixes and the type of the programming language?
 
@@
 
-It is interesting that majority of these languages are not used in Testing, which contains only Java, C, and C++.
-Similarity, only a few languages appears in Code Generators and Site Management categories.
+It is interesting that the majority of these languages are not used in Testing, which contains only Java, C, and C++.
+Similarity, only a few languages appear in Code Generators and Site Management categories.
 
@@
 They are expected to be tested extensively before deployment to reduce the chance of failure.
-This may results in less hot fixes being needed after deployment.
+This may result in fewer hotfixes being needed after deployment.
 An interesting observation is that the average hotfix ratio of C in Internet category is incredibly low (0.03).
-We believe that is because of there are several new languages designed specifically for Internet applications.
-Not many new C code are being written in this area which results in a very low hotfix commits.
+We believe that is because there are several new languages designed specifically for Internet applications.
+Not much new C code is being written in this area, which results in a very low number of hotfix commits.
 %Security categories  %This could be down to fully comprehensive testing techniques, but could also be attributed to the fact that within those categories, the language has been used for an extensive period and hence there is a lot of experience with a variety of different bugs, meaning they are less likely to occur during development as experienced programmers are less likely to make those particular mistakes in the first place.
@@
 
-On the other hand, the categories with \textbf{the highest average hotfix ratios are Code Generators (0.53)}.
+On the other hand, the category with \textbf{the highest average hotfix ratio is Code Generators (0.53)}.
 C and C++, have very high average hot fix ratios while  PHP and Java are the lowest.
@@
 
-The second place is Role playing category (0.47) which has higher hotfix ratio than its superset, Game category.
+The second least reliable category is Role playing (0.47) which has higher hotfix ratio than its superset, Game category.
 We believe that this specific kind of game is more complex with vast amount of configurations of characters and items in the game.
@@
 \subsection{RQ3: hotfixes vs. languages}
-Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has the least average hotfix ratio (0.176)} followed by JavaScript (0.209) and Perl (0.218).
-However, all the mentioned languages show marginally better results with approximately 8\% difference comparing to other common languages: PHP (0.282), C++ (0.286), C (0.305) and C\#(0.339).
+Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has the lowest average hotfix ratio (0.176)} followed by JavaScript (0.209) and Perl (0.218).
+However, all the mentioned languages show marginally better results with an approximately 8\%-difference compared to other common languages: PHP (0.282), C++ (0.286), C (0.305) and C\#(0.339).
 \textbf{C\# has the highest average hotfix ratio among all the languages.}
@@
 \subsection{RQ4: hotfixes vs. language types}
-Another interesting research question is comparing the type of each programming language by their average hotfix ratio.
+Another interesting research question is comparing the type of programming languages by their average hotfix ratios.
 The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared using hotfix ratios with respect to their three main classes.
@@
 Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains  \cite{Wadler1999}.
-Consequently, we only compare procedural and scripting languages in this class.
+Consequently, we only compared procedural and scripting languages in this class.
 
 \subsubsection*{Programming paradigm: procedural vs. scripting}
-From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among procedural (Java, C, C\#, C++) and scripting (Perl, Python, PHP) across all the categories.
-This is possibly because the number of data points (languages) were too low to observe such behaviours and make any firm conclusions.
+From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among procedural (Java, C, C\#, C++) and scripting (Perl, Python, PHP) languages across all the categories.
+This is possibly because the number of data points (languages) were too low to observe such behaviours and draw any firm conclusions.
 %It was always the case that at least one of the languages would not follow a pattern.
@@
 \subsubsection*{Compilation: static vs. dynamic}
-In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}.
+In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear difference of observed hotfix ratios between the two sets as can be seen from Table \ref{tab:1}.
 The reason was because the languages in these two categories coincided with the languages in the first two classes (procedural and scripting).
@@
 Similarly to the previous two classes, we encountered no clear pattern across the categories.
-Furthermore, we found the same problem of having comparisons against only one member of a language class.
-Solely in the Communications category, we were able to detect some pattern, where strongly-typed languages performed slightly better.
+Furthermore, we found problems of having comparisons against only one member of a language class.
+Solely in the Communications category, we were able to detect some pattern, where weakly-typed languages performed slightly better.
 However, there were only two languages labelled as weak.
@@
 
-There is a low amount of languages satisfied our predicate (being present in at least one half of all evaluated topics), which did not allow us to state any conclusive results.
-In addition, with a limited number of programming topics was used (16 in total), we are restricted to observe distinguishable patterns.
+There was a low amount of languages that satisfied our predicate (being present in at least one half of all evaluated topics), which did not allow us to state any conclusive results.
+In addition, with a limited number of programming topics used (16 in total), we are restricted to observe distinguishable patterns.
 
-Lastly, only one dataset was used to evaluate our hypothesis: SourceForge 2013. Hence, this results may not be generalised.
+Lastly, only one dataset was used to evaluate our hypothesis: SourceForge September 2013. Hence, these results may not be generalised.
 
 \section{Conclusions and Future Work}
-We have presented a study in detecting the fault-proneness of programming languages through hotfix commits, a topic that has not been investigated so far.
-As part of our research we performed a topic-specific analysis across 16 programming domains (categories) to detect the most popular languages used in the evaluated  projects.
-We investigated languages and topics that achieved highest/lowest ratios of hotfix per version update and tried to identify patterns based on the language classification presented in \cite{Ray2014}.
-We used Boa to mine repositories in the SourceForge 2013's dataset.
+We have presented a study on detecting the fault-proneness of programming languages through hotfix commits, a topic that has not been investigated so far.
+As part of our research we performed a topic-specific analysis across 16 programming domains (categories) to detect the most widely used languages.
+Languages and topics that achieved highest/lowest ratios of hotfix per version update were investigated and we tried to identify patterns based on the language classification presented in \cite{Ray2014}.
+We used Boa to mine repositories in the SourceForge September 2013's dataset.
 %Our goal was to verify whether the general relationships between languages' reliability holds with regard to hotfixes.
-We identified 8 languages as being the most widely used across different domains including Java, C, C++, Perl, Python, PHP, C\#, and JavaScript respectively.
-We discovered that Java achieves the least number of hotfixes per version update (0.176), while C\# requiring the highest amount of hotfixes per version increase (0.339).
-From the different categories, we observed that Security, Internet and Communications is the most reliable (i.e.
-requiring least amount of hotfixes per version update) while Role Playing and Code Generation required the highest number of corrections by hotfixes.
+We identified 8 languages as being the most popular across different domains including: Java, C, C++, Perl, Python, PHP, C\#, and JavaScript.
+We discovered that Java achieves the smallest hotfix/version update ratio (0.176), while C\# required the highest amount of hotfixes per version increase (0.339).
+From the different categories, we observed that Security, Internet and Communications are the most reliable (i.e.
+achieving the lowest hotfix/version update ratios), while Role Playing and Code Generation experienced the highest number of corrections by hotfixes.
 Finally, according to the language classes, we were not able to identify any conclusive patterns.
 
-We believe that there are still some aspects to be improved in our study.
-In particular, more datasets could be used, including Github (provided that projects are categorised into topics), and more reliable and accurate method to identify bump commits.
-Instead of using data from SourceForge, real commercial projects which have been open-sourced can be investigated to enhance the quality of our data.
+We believe that there are some aspects to be improved in our study.
+In particular, more datasets could be used, including Github (provided that projects are categorised into topics), and more reliable method to identify bump commits.
+Instead of using data from SourceForge, real commercial projects which have been open-sourced can be mined to enhance the quality of our data.
 In such projects, releases should occur more frequently and thus, there is a higher probability of identifying more hotfixes.
 
-Finally, we hope that our paper lays a good foundation for future research to be done in this area.
-We expect that the results we obtained will help developers and project managers improve their planning and understand what are the most appropriate languages for their specific topic or purpose.
+Finally, we hope that our paper lays a good foundation for future research in this area.
+We expect that the results we obtained will help developers and project managers improve their planning and use the most appropriate languages for a specific topic or purpose.
 

### 2016-02-22 msr_paper/main.tex
Update to table and minor corrections.

@@
 
-Our work was inspired by a research paper \cite{Ray2014}, where authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.
-Our contribution primarily involved investigation of the fault-proneness of different languages based on the newly defined criteria - hotfixes per version update.
+Our work was inspired by a previous research paper~\cite{Ray2014} which compared the number of bug fix commits to the total number of commits across 17 languages and 850 different projects.
+Our primary contribution is the investigation of the fault-proneness of different languages accross project domains based on the newly defined criteria \emph{hotfixes per version update}.
 
 \subsection{Dataset}
-The Boa September 2013 SourceFourge dataset is selected for this study.
-It consists of 49,865 projects, selected from the top 16 most popular project categories in the dataset.
-We only concentrate on projects that incorporate versioning (i.e.
-containing bump commits).
-%That enables us to differentiate results of each different project category.
+We selected the Boa September 2013 SourceFourge dataset for this study as it containts project topics.
+From the dataset, we selected projects from the top 16 most popular project topics.
+This resulted in of 49,865 projects.
+Moreover, we only concentrate on projects that incorporate versioning (i.e.\ containing bump commits).
 
 \subsection{Approach for identifying hotfix commits}
-We invented a method to distinguish a hotfix based on commit messages. The method has 2 parts:
+Our method to distinguish hotfixes based on commit messages has 2 parts:
 
@@
 
-\noindent\textbf{RQ2 (hotfixes vs. project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ2 (hotfixes vs.\ project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ3 (hotfixes vs. languages):} What languages achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ3 (hotfixes vs.\ languages):} What languages achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ4 (hotfixes vs. language types):} Are there any relationships between the occurrences of hotfixes and the type of the programming language?
+\noindent\textbf{RQ4 (hotfixes vs.\ language types):} Are there any relationships between the occurrences of hotfixes and the type of the programming language?
 
@@
 	\small
-	\begin{tabular}{l c c c r r r r r r r r r r r r r r r r r}
+	\begin{tabular}{@{}l c@{~~}c@{~~}c r r r r r r r r r r r r r r r r r@{}}
 		Language & $\mathcal{P}$ & $\mathcal{C}$ & $\mathcal{T}$ & 
@@
 
-\section{Results}
+%\section{Results}
 We performed an empirical study on the SourceForge dataset using our method of identifying version bump commits and hotfix commits. The findings are explained below.
@@
 
-\subsection{RQ2: hotfixes vs. project categories}
+\subsection{RQ2: hotfixes vs.\ project categories}
 To be able to compare among languages, we calculated their hotfix ratios.
@@
 
-\subsection{RQ3: hotfixes vs. languages}
+\subsection{RQ3: hotfixes vs.\ languages}
 Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has the lowest average hotfix ratio (0.176)} followed by JavaScript (0.209) and Perl (0.218).
@@
 
-\subsection{RQ4: hotfixes vs. language types}
+\subsection{RQ4: hotfixes vs.\ language types}
 Another interesting research question is comparing the type of programming languages by their average hotfix ratios.
@@
 
-\subsubsection*{Programming paradigm: procedural vs. scripting}
+\subsubsection*{Programming paradigm: procedural vs.\ scripting}
 From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among procedural (Java, C, C\#, C++) and scripting (Perl, Python, PHP) languages across all the categories.
@@
 
-\subsubsection*{Compilation: static vs. dynamic}
+\subsubsection*{Compilation: static vs.\ dynamic}
 In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear difference of observed hotfix ratios between the two sets as can be seen from Table \ref{tab:1}.
@@
 
-\subsubsection*{Type: strongly-type vs. weakly-typed}
+\subsubsection*{Type: strongly-type vs.\ weakly-typed}
 The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.

### 2016-02-22 msr_paper/main.tex
Some more fixes.

@@
 We analyse the prevalence of hotfix commits in various development areas as an indicator of which languages tend to produce the least amount of problems after software has been published.
-Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are used across project types.
+Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project topics and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are used across project domains.
 Our results only identified 8 languages that are widely used across different domains: Java, C, C++, C\#, Perl, Python, PHP, and JavaScript.
 Java has the least number of hotfixes per version update (0.18), while C\# required most hotfixes per version update (0.34).
-We identified the Security and Internet \& Communications project types as being the most reliable (i.e.\ requiring least amount of hotfixes per version update).
+We identified the Security and Internet \& Communications project topic as being the most reliable (i.e.\ requiring least amount of hotfixes per version update).
 The analysis did not reveal conclusive patterns for language classes (strongly vs.\ weakly typed and dynamic vs.\ static).
@@
   \item A quantitative analysis identifying fault-proneness of specific programming topics and the distribution of faults across programming languages.
-The analysis identified that only eight languages are used across the top 16 most popular domains identified by the Boa SourceForge 2013 dataset.
+The analysis identified that only eight languages are used across the top 16 most popular domains (topics) identified by the Boa SourceForge 2013 dataset.
   \item An investigation on the impact of language classes on reliability: strongly-typed vs weakly-typed, static vs dynamic and scripting vs procedural.
@@
 
-\subsection{Comparison of programming languages}
+\subsection{Comparison of Programming Languages}
 %Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality \cite{Agrawal2015}.
@@
 The study was carried out on a huge dataset from GitHub (1.5 million commits).
-To evaluate the effect that programming languages have on software quality, the authors grouped languages according to their type, e.g.\ strongly vs.\ weakly typed.
+To evaluate the effect that programming languages have on software quality, the authors grouped languages according to their class, e.g.\ strongly vs.\ weakly typed.
 They also used mathematical methods to triangulate and produce statistics for how group size, project history and number of contributors affect the number of defects.
@@
 
-Other work used different parameters to compare programming languages, for example, program length, programming effort, run-time efficiency, memory consumption, and reliability \cite{Prechelt2000}.
+Other work used different parameters to compare programming languages, for example, program length, programming effort, run-time efficiency, memory consumption, and reliability~\cite{Prechelt2000}.
 Fourment and Gilling~\cite{Fourment2008} analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages.
@@
 
-\subsection{Identifying bug fix commits}
+\subsection{Identifying Bug Fix Commits}
 %The first major problem we had to solve was identifying actual hotfix commits.
@@
 Our work was inspired by a previous research paper~\cite{Ray2014} which compared the number of bug fix commits to the total number of commits across 17 languages and 850 different projects.
-Our primary contribution is the investigation of the fault-proneness of different languages accross project domains based on the newly defined criteria \emph{hotfixes per version update}.
+Our primary contribution is the investigation of the fault-proneness of different languages \emph{across} project domains (topics) based on the newly defined criteria \emph{hotfixes per version update}.
 
 \subsection{Dataset}
-We selected the Boa September 2013 SourceFourge dataset for this study as it containts project topics.
+We selected the Boa September 2013 SourceFourge dataset for this study as it contains project topics.
 From the dataset, we selected projects from the top 16 most popular project topics.
@@
 
-\subsection{Approach for identifying hotfix commits}
-Our method to distinguish hotfixes based on commit messages has 2 parts:
+\subsection{Approach for Identifying Hotfix Commits}
+Our method to distinguish hotfixes based on commit messages has two parts:
 
@@
 
-\subsection{Research questions}
+\subsection{Research Questions}
 \noindent Our study aimed to answer the following research questions:
 
-\noindent\textbf{RQ1 (language popularity):} What languages are the most widely used across the top 16 most popular categories?
+\noindent\textbf{RQ1 (language popularity):} Which languages are the most widely used across the top 16 most popular topics?
 
-\noindent\textbf{RQ2 (hotfixes vs.\ project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ2 (hotfixes vs.\ project topic):} Which project topics achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ3 (hotfixes vs.\ languages):} What languages achieve the highest/lowest number of hotfixes per version update?
+\noindent\textbf{RQ3 (hotfixes vs.\ languages):} Which languages achieve the highest/lowest number of hotfixes per version update?
 
-\noindent\textbf{RQ4 (hotfixes vs.\ language types):} Are there any relationships between the occurrences of hotfixes and the type of the programming language?
+\noindent\textbf{RQ4 (hotfixes vs.\ language class):} Are there any relationships between the occurrences of hotfixes and the class of the programming language?
 
@@
 	\end{tabular}
-	\caption{Ratios of hotfix commits per version bump commits of 8 common languages and their 16 respective categories.
+	\caption{Ratios of hotfix commits per version bump commits of 8 common languages and their 16 respective topics.
 $\mathcal{P}$, $\mathcal{C}$, and $\mathcal{T}$ represent the classes of the language.
 $\mathcal{P}$ stands for Programming Paradigm class (P=Procedural, S=Scripting).
-$\mathcal{C}$ is for Compilation class (S=Static, D=Dynamic), and $\mathcal{T}$ is for Type class (W=weakly-typed, S=Strongly-typed).}
+$\mathcal{C}$ is for the Compilation class (S=Static, D=Dynamic), and $\mathcal{T}$ is for the Type class (W=weakly-typed, S=Strongly-typed).}
 	\label{tab:1}
@@
 We performed an empirical study on the SourceForge dataset using our method of identifying version bump commits and hotfix commits. The findings are explained below.
-\subsection{RQ1: language popularity}
-Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages.
-We chose the languages based on their frequency of occurrence across 16 categories.
+\subsection{RQ1: Language Popularity}
+Our investigation on the most widely used languages across the top 16 most popular topics revealed that there are surprisingly few such languages.
+We chose the languages based on their frequency of occurrence across 16 topics.
 From Table \ref{tab:1}, the 8 widely used languages include Java (16), C (15), C++ (15), Perl (11), Python (11), PHP (11), C\# (10), and JavaScript (7) respectively.
@@
 It is interesting that the majority of these languages are not used in Testing, which contains only Java, C, and C++.
-Similarity, only a few languages appear in Code Generators and Site Management categories.
+Similarity, only a few languages appear in the Code Generators and the Site Management topics.
 
@@
 
-\subsection{RQ2: hotfixes vs.\ project categories}
+\subsection{RQ2: Hotfixes vs.\ Project Topics}
 To be able to compare among languages, we calculated their hotfix ratios.
-The hotfix ratio of a language $P$ in category $C$ is a ratio of total hotfix commits per total version bump commits of the language in that category.
+The hotfix ratio of a language $P$ in topic $C$ is a ratio of total hotfix commits per total version bump commits of the language in that topic.
 %We proceed to analyse why certain categories might have low numbers of hotfixes and vice versa.
@@
 \end{equation*}
-In Table \ref{tab:1}, \textbf{Communications have the lowest average hotfix ratios across all categories (0.11)} followed by Internet, Security, and Systems Administration (0.12, 0.17, 0.17) respectively.
-These categories are information critical.
+In Table \ref{tab:1}, \textbf{Communications have the lowest average hotfix ratios across all topics (0.11)} followed by Internet, Security, and Systems Administration (0.12, 0.17, 0.17) respectively.
+These topics are information critical.
 They are expected to be tested extensively before deployment to reduce the chance of failure.
 This may result in fewer hotfixes being needed after deployment.
-An interesting observation is that the average hotfix ratio of C in Internet category is incredibly low (0.03).
+An interesting observation is that the average hotfix ratio of C in the Internet topic is incredibly low (0.03).
 We believe that is because there are several new languages designed specifically for Internet applications.
@@
 
-On the other hand, the category with \textbf{the highest average hotfix ratio is Code Generators (0.53)}.
-C and C++, have very high average hot fix ratios while  PHP and Java are the lowest.
+On the other hand, the topic with \textbf{the highest average hotfix ratio is Code Generators (0.53)}.
+In this topic, C and C++ have very high average hot fix ratios while  PHP and Java have the lowest.
 One can speculate that PHP and Java are established languages for writing code generators and due to this engender low hot fix averages.
@@
 
-The second least reliable category is Role playing (0.47) which has higher hotfix ratio than its superset, Game category.
+The second least reliable topic is Role playing (0.47) which has higher hotfix ratio than its superset, the Game topic.
 We believe that this specific kind of game is more complex with vast amount of configurations of characters and items in the game.
@@
 
-\subsection{RQ3: hotfixes vs.\ languages}
-Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has the lowest average hotfix ratio (0.176)} followed by JavaScript (0.209) and Perl (0.218).
-However, all the mentioned languages show marginally better results with an approximately 8\%-difference compared to other common languages: PHP (0.282), C++ (0.286), C (0.305) and C\#(0.339).
+\subsection{RQ3: Hotfixes vs.\ Languages}
+The language comparison based on the average hotfix ratio over all the topics in Table 1 has shown that \textbf{Java has the lowest average hotfix ratio (0.176)} followed by JavaScript (0.209) and Perl (0.218).
+However, all the mentioned languages show marginally better results compared to other common languages: PHP (0.282), C++ (0.286), C (0.305) and C\#(0.339).
 \textbf{C\# has the highest average hotfix ratio among all the languages.}
 
-\subsection{RQ4: hotfixes vs.\ language types}
-Another interesting research question is comparing the type of programming languages by their average hotfix ratios.
-The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared using hotfix ratios with respect to their three main classes.
-Programming Paradigm class (noted as $\mathcal{P}$ in Table \ref{tab:1}) has three variations: procedural, functional, and scripting.
-The Compilation class contains static and dynamic type.
-The final one, Type class, separates languages into strongly-typed and weakly-typed.
-According to \cite{Ray2014}, strongly-typed languages experienced fewer bug fixes than its counterpart, weakly-typed.
+\subsection{RQ4: Hotfixes vs.\ Language Classes}
+Another interesting research question is comparing the class of programming languages by their average hotfix ratios.
+The classification in Table \ref{tab:1} is inspired by Ray et al.~\cite{Ray2014}, where languages were compared using hotfix ratios with respect to their three main classes.
+The Programming Paradigm class (noted as $\mathcal{P}$ in Table \ref{tab:1}) has three variations: procedural, functional, and scripting.
+The Compilation class contains static and dynamic typing.
+The final Type class separates languages into strongly-typed and weakly-typed.
+According to Ray et al.~\cite{Ray2014}, strongly-typed languages experienced fewer bug fixes than its counterpart, weakly-typed.
 Static programming languages required fewer corrective commits than dynamic ones.
@@
 Due to our criteria of selecting the languages, the chosen 8 languages are all either procedural or scripting, but not functional.
-Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains  \cite{Wadler1999}.
+Most probably this was due to the fact that functional languages are limited to specific use cases and as such are not widely popular across different domains~\cite{Wadler1999}.
 Consequently, we only compared procedural and scripting languages in this class.
 
-\subsubsection*{Programming paradigm: procedural vs.\ scripting}
-From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among procedural (Java, C, C\#, C++) and scripting (Perl, Python, PHP) languages across all the categories.
+\subsubsection*{Programming Paradigm: Procedural vs.\ Scripting}
+From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among procedural (Java, C, C\#, C++) and scripting (Perl, Python, PHP) languages across all the topics.
 This is possibly because the number of data points (languages) were too low to observe such behaviours and draw any firm conclusions.
@@
 
-\subsubsection*{Compilation: static vs.\ dynamic}
+\subsubsection*{Compilation: Static vs.\ Dynamic}
 In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear difference of observed hotfix ratios between the two sets as can be seen from Table \ref{tab:1}.
-The reason was because the languages in these two categories coincided with the languages in the first two classes (procedural and scripting).
+The reason was because the languages in these two classes coincided with the languages in the first two classes (procedural and scripting).
 
-\subsubsection*{Type: strongly-type vs.\ weakly-typed}
+\subsubsection*{Type: Strongly-Typed vs.\ Weakly-Typed}
 The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages.
-Similarly to the previous two classes, we encountered no clear pattern across the categories.
+Similarly to the previous two classes, we encountered no clear pattern across the topics.
 Furthermore, we found problems of having comparisons against only one member of a language class.
-Solely in the Communications category, we were able to detect some pattern, where weakly-typed languages performed slightly better.
+Solely in the Communications topic, we were able to detect some pattern, where weakly-typed languages performed slightly better.
 However, there were only two languages labelled as weak.
@@
 We have presented a study on detecting the fault-proneness of programming languages through hotfix commits, a topic that has not been investigated so far.
-As part of our research we performed a topic-specific analysis across 16 programming domains (categories) to detect the most widely used languages.
-Languages and topics that achieved highest/lowest ratios of hotfix per version update were investigated and we tried to identify patterns based on the language classification presented in \cite{Ray2014}.
-We used Boa to mine repositories in the SourceForge September 2013's dataset.
+As part of our research we performed a topic-specific analysis across 16 programming domains (topics) to detect the most widely used languages.
+Languages and topics that achieved highest/lowest ratios of hotfix per version update were investigated and we tried to identify patterns based on the language classification presented by Ray et al.~\cite{Ray2014}.
+We used Boa to mine repositories in the SourceForge September 2013 dataset.
 %Our goal was to verify whether the general relationships between languages' reliability holds with regard to hotfixes.
@@
 We discovered that Java achieves the smallest hotfix/version update ratio (0.176), while C\# required the highest amount of hotfixes per version increase (0.339).
-From the different categories, we observed that Security, Internet and Communications are the most reliable (i.e.
+From the different topics, we observed that Security, Internet and Communications are the most reliable (i.e.
 achieving the lowest hotfix/version update ratios), while Role Playing and Code Generation experienced the highest number of corrections by hotfixes.
